---
title: "Comparison of naive CCL19-EYFP cells vs mCOV-FIt31-g33 CCL19-EYFP cells"
author: "Chrysa Papadopoulou"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: yes
    collapsed: no
    smooth_scroll: yes
mainfont: {Arial Unicode MS}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("ragg_png", "cairo_pdf"))
knitr::asis_output("\U207A")
knitr::spin_child('extra_functions.R')
set.seed(1000)
```

### Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(here)
  library(purrr)
  library(dplyr)
  library(stringr)
  library(patchwork)
  library(Seurat)
  library(Matrix)
  library(gridExtra)
  library(gsubfn)
  library(ggsci)
  library(biomaRt)
  library(tidyverse)
  library(msigdbr)
  library(stats)
  library(clusterProfiler)
  library(dict)
  library(openxlsx)
  library(DOSE)
  library(enrichplot)
})
```

### Set directory
```{r set directory}
basedir <- here()
```

### Read CCL19-EYFP`r knitr::asis_output("\U207A")` cell data from naïve lungs and excised LLC-gp33 tumors on day 23
```{r read CCL19 FRC data}
CCL19_EYFP <- readRDS(paste0(basedir,"/data/Mouse/CCL19_EYFP_nonmCOV.rds"))
```

### Read CCL19-EYFP`r knitr::asis_output("\U207A")` mCOV-FIt31-g33 cell data
```{r read CCL19 FRC mCOV-FIt31-g33 data}
CCL19_EYFP_mCOV <- readRDS(paste0(basedir,"/data/Mouse/mCOV.rds"))
```

### Feature plots {.tabset}
#### Cxcl13 LLC naive
```{r CXCL13 LLC naive}
FeaturePlot(CCL19_EYFP, reduction = "umap",
          features = get_full_gene_name('Cxcl13',CCL19_EYFP),raster=FALSE,
          cols=c("lightgrey", "darkred"),min.cutoff = 0, max.cutoff = 6) + ggtitle(paste0("Ccl19-EYFP", "\U207A ", "cells (Naive)"))
```

#### Cxcl13 mCOV-FIt31-g33
```{r CXCL13 mCOV}
FeaturePlot(CCL19_EYFP_mCOV, reduction = "umap", 
          features = get_full_gene_name('Cxcl13',CCL19_EYFP_mCOV),raster=FALSE,
          cols=c("lightgrey", "darkred")) + ggtitle(paste0("Ccl19-EYFP", "\U207A ", "cells (mCOV-FIt31-g33)"))
```

#### Cr2 mCOV-FIt31-g33
```{r CR2 mCOV}
FeaturePlot(CCL19_EYFP_mCOV, reduction = "umap", 
          features = get_full_gene_name('Cr2',CCL19_EYFP_mCOV),raster=FALSE,
          cols=c("lightgrey", "darkred")) + ggtitle(paste0("Ccl19-EYFP", "\U207A ", "cells (mCOV-FIt31-g33)"))
```

**Let's compare the gene expression between TRC/PRC (d23) and their progenitor subsets (d15) to identify gene programs that may be relevant for their function in supporting the T cell niches**

```{r progenitors d15 and d23}
#Set new annotation: Change Smoc1 AdvFB to Cd34 AdvFB
CCL19_EYFP@meta.data$annot[CCL19_EYFP@meta.data$annot == paste0("Smoc1", expression("\u207A "), "AdvFB") ] <- paste0("Cd34", expression("\u207A "), "AdvFB")

progenitors_d15 <- subset(CCL19_EYFP, TimePoint == "d15" & annot %in% c(paste0("Cd34", expression("\u207A "), "AdvFB"),"SMC/PC"))

progenitors_mCOV <- subset(CCL19_EYFP_mCOV, annot %in% c(paste0("Rgs5", expression("\u207A "), "PRC"),paste0("Sulf1", expression("\u207A "), "TRC"),"TLS TRC"))

data_merge <- merge(progenitors_d15, y = c(progenitors_mCOV),
             add.cell.ids = c("progenitors_d15","progenitors_mCOV"),
             project = "progenitors_d15_mCOV")

#Preprocessing
resolution <- c(0.1, 0.25, 0.4, 0.6,0.8, 1.)
data_merge  <- preprocessing(data_merge,resolution)
```

### Save merged data progenitors Day 15 (naive) and Day 23 (mCOV-FIt31-g33)
```{r save progenit d15 d23}
#saveRDS(data_merge,paste0(basedir,"/data/Human/progenitors_d15_mCOV.rds"))
```

### Pathway analysis{.tabset}
#### **CD34`r knitr::asis_output("\U207A")` AdvFB (Day 15) vs TLS TRC (Day 23)** (Supplementary Figure 5I)
```{r AdvFB_TLS}
# Step 1 : Set output directory
subDir <- "GSEA_AdvFB_TLS/"
saving_path <- paste0(basedir,"/output/")
final_dir <- file.path(saving_path, subDir)
dir.create(final_dir, showWarnings = FALSE,recursive = TRUE)
map_df <- ExtractMouseGeneSets(final_dir)

# Step 2: Customize parameters
httr::set_config(httr::config(ssl_verifypeer = FALSE))
organism <- "org.Mm.eg.db"

disease_phase <- "d15vsd23_AdvFB_TLS"
datatype <- "SYMBOL"

advfb_tls <-subset(data_merge, annot %in% c(paste0("Cd34", expression("\u207A "), "AdvFB"),"TLS TRC"))

Idents(advfb_tls) <- advfb_tls$annot
  
DEmarkers <-FindAllMarkers(advfb_tls, only.pos=T, logfc.threshold = 0.1,
                           min.pct = 0.1)

Vec <-unique(advfb_tls$annot)
EnrichParameters_TLS <-customize_parameters(Vec,DEmarkers,organism,datatype,disease_phase,saving_path) 

# Step 3: Enrichment Analysis
for (i in seq(1,length(EnrichParameters_TLS$enrichcl_list))){
  terms<- EnrichParameters_TLS$enrichcl_list[[i]]
  # Filter on the most significant pathways (keep rows where p.adjust<= 0.05)
  terms<- terms@result[terms@result$p.adjust <= 0.05,]
  population <- Vec[i]
  population<- gsub("/", "_", population)
  write.xlsx(terms, paste0(final_dir,"/","GO_Pathways_",population,".xlsx"),row.names = TRUE)
}

#Step 4: Plot enriched pathways
pathways <-c("cell chemotaxis", "tissue remodeling","activation of immune response",
             "leukocyte mediated immunity", "interleukin-1 production", "tumor necrosis factor production")

TLS_terms <- EnrichParameters_TLS$enrichcl_list[[2]]@result

selec_pathways <- TLS_terms[TLS_terms$Description %in% pathways,]

selec_pathways$Description <- factor(selec_pathways$Description, levels = rev(pathways))
selec_pathways <- selec_pathways[order(selec_pathways$Description), ]

ggplot(data=selec_pathways, aes(x=Description, y=qscore, fill = analysis)) + xlab(NULL) +
    geom_bar(stat="identity",position="dodge",colour = "black",show.legend = FALSE, width= 0.8, size = 1 ) + coord_flip() +
    scale_y_continuous(expand = expansion(c(0,0)), limits = c(0.0, 6),breaks = c(0,2,4,6)) +
    scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 80)) +
    theme( legend.justification = "top",
           plot.title = element_text(hjust = 0.5,size = 12,face="bold"),axis.line = element_line(colour = "black"),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle = 0, vjust = 0.5,colour = "black", size = 10),
           axis.text.y = element_text(angle = 0, vjust = 0.8,colour = "black", size = 10, face=ifelse(levels(selec_pathways$Description)=="tumor necrosis factor production","bold","plain")),
           axis.title.y = element_text(size = rel(2), angle = 45),
           axis.title.x = element_text(size = rel(1.5), angle = 0),
           axis.text = element_text(size = 8),
           panel.background = element_blank(), legend.position = "none") +
    scale_fill_manual(values = "dark gray") + ggtitle("Enriched in TLS TRC (vs. AdvFB day 15)")
```

#### **CD34`r knitr::asis_output("\U207A")` AdvFB (Day 15) vs Sulf1`r knitr::asis_output("\U207A")` TRC (Day 23)** (Supplementary Figure 5J)
```{r AdvFB_SULF1_TRC}
# Step 1 : Set output directory
subDir <- "GSEA_AdvFB_SULF1/"
saving_path <- paste0(basedir,"/output/")
final_dir <- file.path(saving_path, subDir)
dir.create(final_dir, showWarnings = FALSE,recursive = TRUE)
map_df <- ExtractMouseGeneSets(final_dir)

# Step 2: Customize parameters
httr::set_config(httr::config(ssl_verifypeer = FALSE))
organism <- "org.Mm.eg.db"

disease_phase <- "d15vsd23_AdvFB_SULF1"
datatype <- "SYMBOL"

AdvFB_sulf1 <-subset(data_merge, annot %in% c(paste0("Cd34", expression("\u207A "), "AdvFB"),paste0("Sulf1", expression("\u207A "), "TRC")))

Idents(AdvFB_sulf1) <- AdvFB_sulf1$annot
  
DEmarkers <-FindAllMarkers(AdvFB_sulf1, only.pos=T, logfc.threshold = 0.1,
                           min.pct = 0.1)

Vec <-unique(AdvFB_sulf1$annot)
EnrichParameters_Sulf1 <-customize_parameters(Vec,DEmarkers,organism,datatype,disease_phase,saving_path) 

# Step 3: Enrichment Analysis
for (i in seq(1,length(EnrichParameters_Sulf1$enrichcl_list))){
  terms<- EnrichParameters_Sulf1$enrichcl_list[[i]]
  # Filter on the most significant pathways (keep rows where p.adjust<= 0.05)
  terms<- terms@result[terms@result$p.adjust <= 0.05,]
  population <- Vec[i]
  population<- gsub("/", "_", population)
  write.xlsx(terms, paste0(final_dir,"/","GO_Pathways_",population,".xlsx"),row.names = TRUE)
}

#Step 4: Plot enriched pathways
pathways <-c("regulation of cytokine-mediated signaling pathway", "lymphocyte homeostasis","actin filament bundle organization",
             "positive regulation of cell-cell adhesion", "regulation of T cell activation", "tissue remodeling")

TRC_term_sulf1 <- EnrichParameters_Sulf1$enrichcl_list[[2]]@result

selec_pathways <- TRC_term_sulf1[TRC_term_sulf1$Description %in% pathways,]

selec_pathways$Description <- factor(selec_pathways$Description, levels = rev(pathways))
selec_pathways <- selec_pathways[order(selec_pathways$Description), ]

ggplot(data=selec_pathways, aes(x=Description, y=qscore, fill = analysis)) + xlab(NULL) +
    geom_bar(stat="identity",position="dodge",colour = "black",show.legend = FALSE, width= 0.8, size = 1 ) + coord_flip() +
    scale_y_continuous(expand = expansion(c(0,0)), limits = c(0.0, 6),breaks = c(0,2,4,6)) +
    scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 80)) +
    theme( legend.justification = "top",
           plot.title = element_text(hjust = 0.5,size = 12,face="bold"),axis.line = element_line(colour = "black"),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle = 0, vjust = 0.5,colour = "black", size = 10),
           axis.text.y = element_text(angle = 0, vjust = 0.8,colour = "black", size = 10, face=ifelse(levels(selec_pathways$Description)=="regulation of T cell activation","bold","plain")),
           axis.title.y = element_text(size = rel(2), angle = 45),
           axis.title.x = element_text(size = rel(1.5), angle = 0),
           axis.text = element_text(size = 8),
           panel.background = element_blank(), legend.position = "none") +
    scale_fill_manual(values = "dark gray") + ggtitle(paste0("Enriched in Sulf1", "\U207A ", "TRC (vs. AdvFB day 15)")) 
```

#### **SMC/PC (Day 15) vs PRC (Day 23)** (Supplementary Figure 5K)
```{r SMC_PRC}
# Step 1 : Set output directory
subDir <- "GSEA_SMC_PRC/"
saving_path <- paste0(basedir,"/output/")
final_dir <- file.path(saving_path, subDir)
dir.create(final_dir, showWarnings = FALSE,recursive = TRUE)
map_df <- ExtractMouseGeneSets(saving_path)

# Step 2: Customize parameters
httr::set_config(httr::config(ssl_verifypeer = FALSE))
organism <- "org.Mm.eg.db"

disease_phase <- "d15vsd23_SMC_PRC"
datatype <- "SYMBOL"

smc_prc <-subset(data_merge, annot %in% c("SMC/PC", paste0("Rgs5", expression("\u207A "), "PRC")))

Idents(smc_prc) <- smc_prc$annot
  
DEmarkers <-FindAllMarkers(smc_prc, only.pos=T, logfc.threshold = 0.1,
                           min.pct = 0.1)

Vec <-unique(smc_prc$annot)
EnrichParameters <-customize_parameters(Vec,DEmarkers,organism,datatype,disease_phase,saving_path) 

# Step 3: Enrichment Analysis
for (i in seq(1,length(EnrichParameters$enrichcl_list))){
  terms<- EnrichParameters$enrichcl_list[[i]]
  # Filter on the most significant pathways (keep rows where p.adjust<= 0.05)
  terms<- terms@result[terms@result$p.adjust <= 0.05,]
  population <- Vec[i]
  population<- gsub("/", "_", population)
  write.xlsx(terms, paste0(final_dir,"/","GO_Pathways_",population,".xlsx"),row.names = TRUE)
}

# Step 4: Plot enriched pathways 
pathways <-c("T cell proliferation", "leukocyte migration","T cell mediated immunity",
             "response to type II interferon", "regulation of angiogenesis", "cytokine-mediated signaling pathway")

PRC_terms <- EnrichParameters$enrichcl_list[[2]]@result

selec_pathways <- PRC_terms[PRC_terms$Description %in% pathways,]

selec_pathways$Description <- factor(selec_pathways$Description, levels = rev(pathways))
selec_pathways <- selec_pathways[order(selec_pathways$Description), ]

ggplot(data=selec_pathways, aes(x=Description, y=qscore, fill = analysis)) + xlab(NULL) +
    geom_bar(stat="identity",position="dodge",colour = "black",show.legend = FALSE, width= 0.8, size = 1 ) + coord_flip() +
    scale_y_continuous(expand = expansion(c(0,0)), limits = c(0.0, 15),breaks = c(0,5,10,15)) +
    scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 80)) +
    theme( legend.justification = "top",
           plot.title = element_text(hjust = 0.5,size = 12,face="bold"),axis.line = element_line(colour = "black"),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle = 0, vjust = 0.5,colour = "black", size = 10),
           axis.text.y = element_text(angle = 0, vjust = 0.8,colour = "black", size = 10, face=ifelse(levels(selec_pathways$Description)=="cytokine-mediated signaling pathway","bold","plain")),
           axis.title.y = element_text(size = rel(2), angle = 45),
           axis.title.x = element_text(size = rel(1.5), angle = 0),
           axis.text = element_text(size = 8),
           panel.background = element_blank(), legend.position = "none") +
    scale_fill_manual(values = "dark gray") + ggtitle("Enriched in PRC (vs. SMC/PC day 15)")
```

### Gene-Concept Networks {.tabset}
#### Enriched in TLS TRC (vs. AdvFB day 15) (Supplementary Figure 5L)
```{r network plot AdvFB_TLS}
pathways <- c("tumor necrosis factor production")

cnetplot(EnrichParameters_TLS$enrichcl_list[[2]], node_label="gene", layout = "kk", showCategory = pathways,
        max.overlaps=Inf,color.params = list(gene ="black",
                                             category = "red",
                                             edge = TRUE),
                          cex.params = list(category_label = 0.0000001,
                                            label_gene = 0.000001, gene_label= 0.6)) + theme(legend.text=element_text(size=8))
```

#### Enriched in Sulf1`r knitr::asis_output("\U207A")` TRC (vs. AdvFB day 15) (Supplementary Figure 5M)
```{r network plot AdvFB_Sulf1}
pathways <- c("regulation of T cell activation")

cnetplot(EnrichParameters_Sulf1$enrichcl_list[[2]], node_label="gene", layout = "kk", showCategory = pathways,
        max.overlaps=Inf,color.params = list(gene ="black",
                                             category = "red",
                                             edge = TRUE),
                          cex.params = list(category_label = 0.0000001,
                                            label_gene = 0.000001, gene_label= 0.6)) + theme(legend.text=element_text(size=8))
```

### Session info
```{r session info}
sessionInfo()
date()
```


