---
title: "Merge NSCLC FRCs and TILs"
author: "Chrysa Papadopoulou"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
mainfont: {Arial Unicode MS}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("ragg_png", "cairo_pdf"))
knitr::asis_output("\U207A")
knitr::spin_child('extra_functions.R')
set.seed(1000)
```

## Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(here)
  library(purrr)
  library(stringr)
  library(patchwork)
  library(Seurat)
  library(Matrix)
  library(dittoSeq)
  library(gridExtra)
  library(CellChat)
  library(NMF)
  library(pheatmap)
})
```

## **Infiltrating lymphocytes in NSCLC**

### Set directory
```{r set directory}
basedir <- here()
```

## **Merge CCL19`r knitr::asis_output("\U207A")` FRCs and Infiltrating lymphocytes in NSCLC**

### Read NSCLC TIL and FRC data
```{r read dataC}
NSCLS_TIL_data <-readRDS(paste0(basedir,"/data/Human/NSCLC_TILs.rds"))
NSCLC_CCL19_FRCs <-readRDS(paste0(basedir,"/data/Human/NSCLC_CCL19_FRCs_CAFs.rds"))
```

### Merge NSCLS CCL19`r knitr::asis_output("\U207A")` FRCs and NSCLS TILs
```{r merge CCL19 FRCs TILS}
same_columns <- intersect(colnames(NSCLS_TIL_data@meta.data),colnames(NSCLC_CCL19_FRCs@meta.data))

NSCLS_TIL_data@meta.data <-NSCLS_TIL_data@meta.data[,same_columns]
NSCLC_CCL19_FRCs@meta.data <-NSCLC_CCL19_FRCs@meta.data[,same_columns]

merged_data<- merge(NSCLS_TIL_data, y = c(NSCLC_CCL19_FRCs),
             add.cell.ids = c('NSCLS_TIL_data','NSCLC_CCL19_FRCs'),
             project = "NSCLC_FRC_TIL")

resolution <- c(0.1, 0.25, 0.4, 0.6, 0.8, 1.,1.2,1.4,1.6,2.)
merged_data  <- preprocessing(merged_data,resolution)
```

### NSCLS CCL19`r knitr::asis_output("\U207A")` FRCs and NSCLS TILs
```{r umap}
#Define color palet
palet <- c("#1B9E77", "#54B0E4","#E3BE00", "#E41A1C", "#4DAF4A","#FB9A99","#377EB8","#A65628","#222F75")
names(palet) <- c( "CAF2/TRC","CAF1/PRC","AdvFB" ,"SMC/PC", paste0("CD4", "\u207A ", "T cells"), paste0("CD8", "\u207A ", "T cells"), "B cells", "Regulatory T cells",paste0("Cycling CD8", "\u207A ", "T cells"))

DimPlot(merged_data, reduction = "umap", group.by = "cell_type", cols=palet)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
         panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle(paste0("NSCLC CCL19", "\U207A ", "FRCs", " and TILs"))

```

### Save merged data
```{r save data}
#saveRDS(merged_data, paste0(basedir,"/data/Human/NSCLC_TILs_SI4.rds"))
```

### **Interactome Analysis with Cellchat (Suoqin Jin et al., 2021)**

### Convert seurat object to cellchat object
```{r conv seurat to cellchat}
cellchat <- Cellchat_Analysis(merged_data)
cellchat <-CellChatDownstreamAnalysis(cellchat,"human",thresh = 0.05)
```

### Save cellchat data
```{r save cellchat data}
#saveRDS(cellchat,paste0(basedir,"/data/Human/CCL19_FRC_TIL_SIF4_cellchat.rds"))
```

### Interactome analysis (Supplementary Figure 4D)
```{r interactome analysis, message=FALSE }
gg <- netAnalysis_signalingRole_scatter(cellchat,color.use = palet)
gg <- gg + ggtitle("Interactome analysis (Cellchat)")
gg
```

### Outgoing signaling patterns
```{r outgoing signaling}
selectK(cellchat, pattern = "outgoing")
```

### Incoming signaling patterns
```{r incoming signaling}
selectK(cellchat, pattern = "incoming")
```

### Incoming and outgoing patterns
```{r patterns, results='hide', fig.keep = 'none'}
nPatterns <- 7
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing", k = nPatterns, color.use = palet)
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "incoming", k = nPatterns, color.use = palet)
```

### Joint dotplot for incoming and outcoming communication patterns (Supplementary Figure 4E)
```{r joint dotplot in out signaling}

pathways <-c("CXCL","CCL","ICAM","VCAM","MK","FGF","NECTIN","TIGIT","SELL","ITGB2","NOTCH","BAFF","SEMA3","SEMA4","LIGHT","EGF","IFN-II","CD226","IGF","FASLG")

order_list <-c("CAF2/TRC","CAF1/PRC","SMC/PC","AdvFB","Regulatory T cells", "CD4⁺ T cells","CD8⁺ T cells","Cycling CD8⁺ T cells","B cells")

netAnalysis_joint_dot(cellchat,color.use = palet,font.size = 12,pathways = pathways, order_list = order_list)
```

## **Merge CCL19`r knitr::asis_output("\U207A")` FRCs and CD8`r knitr::asis_output("\U207A")` T cell subsets in NSCLC**

### Read CD8`r knitr::asis_output("\U207A")` T cell data
```{r read CD8 T}
NCLS_TIL_FRC_cd8 <-readRDS(paste0(basedir,"/data/Human/NSCLC_TILs_CD8_pop.rds"))
```

### Merge NSCLS CCL19`r knitr::asis_output("\U207A")` FRCs and NSCLS CD8`r knitr::asis_output("\U207A")` T cell subsets
```{r merge NSCLC FRC CD8 T}
same_columns <- intersect(colnames(NCLS_TIL_FRC_cd8@meta.data),colnames(NSCLC_CCL19_FRCs@meta.data))

NCLS_TIL_FRC_cd8@meta.data <-NCLS_TIL_FRC_cd8@meta.data[,same_columns]
NSCLC_CCL19_FRCs@meta.data <-NSCLC_CCL19_FRCs@meta.data[,same_columns]

merged_data<- merge(NCLS_TIL_FRC_cd8, y = c(NSCLC_CCL19_FRCs),
             add.cell.ids = c('NCLS_TIL_FRC_cd8','NSCLC_CCL19_FRCs'),
             project = "NSCLC_FRC_TIL_CD8")

resolution <- c(0.1, 0.25, 0.4, 0.6, 0.8, 1.,1.2,1.4,1.6,2.)
merged_data  <- preprocessing(merged_data,resolution)
```

### NSCLS CCL19`r knitr::asis_output("\U207A")` FRCs and NSCLS CD8`r knitr::asis_output("\U207A")` T cell subsets
```{r umapNSCLS CCL19 FRC CD8 }
#Extend palet for CD8 T cell subsets
palet <- c("#1B9E77", "#54B0E4","#E3BE00", "#E41A1C","#F8766D","#00C08B","#7CAE00","#00B4F0","#F564E3")
names(palet) <- c("CAF2/TRC","CAF1/PRC","AdvFB" ,"SMC/PC",paste0("Stem-like/Naive CD8", "\u207A ", "T cells"),paste0("Exhausted CD8", "\u207A ", "T cells"), paste0("Effector-memory CD8", "\u207A ", "T cells"),paste0("Effector CD8", "\u207A ", "T cells"),paste0("Cycling CD8", "\u207A ", "T cells"))

palet <- palet[names(palet) %in% unique(merged_data$cell_type)]

DimPlot(merged_data, reduction = "umap", group.by = "cell_type", cols = palet)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### NSCLS CCL19`r knitr::asis_output("\U207A")` FRCs and NSCLS CD8`r knitr::asis_output("\U207A")` T cell data 
```{r save NSCLC FRC CD8 T}
#saveRDS(merged_data, paste0(basedir,"/data/Human/NSCLC_TILs_CD8_pop_merge.rds"))
```

### **Interactome Analysis with Cellchat (Suoqin Jin et al., 2021)**

### Convert seurat object to cellchat object
```{r conv seurat to cellchat FRC CD8}
cellchat <- Cellchat_Analysis(merged_data)
cellchat <-CellChatDownstreamAnalysis(cellchat,"human",thresh=0.08)
```

### Save cellchat data
```{r save cellchat data FRC CD8}
#saveRDS(cellchat,paste0(basedir,"/data/Human/CCL19_FRC_TIL_F3E_cellchat.rds"))
```

### Interactome analysis (Figure 4E)
```{r interactome analysis NSCLC FRC CD8, message=FALSE }
gg <- netAnalysis_signalingRole_scatter(cellchat,color.use = palet)
gg <- gg + ggtitle("Interactome analysis (Cellchat)")
gg
```

### Outgoing signaling patterns
```{r outgoing signaling FRC CD8}
selectK(cellchat, pattern = "outgoing")
```

### Incoming signaling patterns
```{r incoming signaling FRC CD8}
selectK(cellchat, pattern = "incoming")
```

### Incoming and outgoing patterns
```{r patterns FRC CD8 T, results='hide', fig.keep = 'none'}
nPatterns <- 8
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing", k = nPatterns, color.use = palet)
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "incoming", k = nPatterns, color.use = palet)
```

### Joint dotplot for incoming and outcoming communication patterns (Figure 3F)
```{r joint dotplot in out signaling FRC CD8}
pathways <-c("MHC-II","LAMININ","FN1","CXCL","CCL","VCAM","ICAM","ITGB2","LIGHT","NECTIN","TIGIT","CD226","EGF","FASLG","SEMA4","IFN-II","IL16","GDF")

order_list <-c("CAF2/TRC","CAF1/PRC",paste0("Stem-like/Naive CD8", "\u207A ", "T cells"),
               paste0("Exhausted CD8", "\u207A ", "T cells"),
               paste0("Effector-memory CD8", "\u207A ", "T cells"),
               paste0("Effector CD8", "\u207A ", "T cells"),
               paste0("Cycling CD8", "\u207A ", "T cells"))

netAnalysis_joint_dot(cellchat,color.use = palet,font.size = 8,pathways = pathways, order_list = order_list, exclude = c("AdvFB" ,"SMC/PC"))
```

### Attraction and adhesion (Figure 3G) {.tabset}
#### Heatmap stroma
```{r heatmap VCAM1 CXCL16 CD34 CXCL12 }
Selgenes <- c("VCAM1","CXCL16", "CD34","CXCL12")
order_list <-c("CAF1/PRC","SMC/PC", "CAF2/TRC","AdvFB")
data_subset <- subset(merged_data, cell_type  %in% order_list)
  
Heatmap(data_subset, Selgenes, order_list = order_list, palet = palet)
```

#### Heatmap CD8`r knitr::asis_output("\U207A")` T cell
```{r heatmap ITGA4 ITGB1 ITGB7 SELL CXCR4}
Selgenes <- c("ITGA4","ITGB1","ITGB7","CXCR6","SELL","CXCR4")
order_list <-c(paste0("Stem-like/Naive CD8", "\u207A ", "T cells"), 
               paste0("Effector-memory CD8", "\u207A ", "T cells"),
               paste0("Effector CD8", "\u207A ", "T cells"),
               paste0("Exhausted CD8", "\u207A ", "T cells"),
               paste0("Cycling CD8", "\u207A ", "T cells") )
data_subset <- subset(merged_data, cell_type  %in% order_list)

Heatmap(data_subset, Selgenes, order_list = order_list, palet = palet)
```

### T cell reactivity (Figure 3H) {.tabset}
#### Heatmap stroma
```{r heatmap NECTIN3 NECTIN2 EGFR FAS IFNGR1 IFNGR2}
Selgenes <- c("NECTIN3","NECTIN2","EGFR","FAS","IFNGR1","IFNGR2")
order_list <-c("CAF1/PRC","SMC/PC", "CAF2/TRC","AdvFB")
data_subset <- subset(merged_data, cell_type  %in% order_list)
  
Heatmap(data_subset, Selgenes, order_list = order_list, palet = palet)
```

#### Heatmap CD8`r knitr::asis_output("\U207A")` T cell
```{r heatmap TIGIT CD226 AREG FASLG IFNG}
Selgenes <- c("TIGIT","CD226","AREG","FASLG","IFNG")
order_list <-c(paste0("Stem-like/Naive CD8", "\u207A ", "T cells"), 
               paste0("Effector-memory CD8", "\u207A ", "T cells"),
               paste0("Effector CD8", "\u207A ", "T cells"),
               paste0("Exhausted CD8", "\u207A ", "T cells"),
               paste0("Cycling CD8", "\u207A ", "T cells") )
data_subset <- subset(merged_data, cell_type  %in% order_list)

Heatmap(data_subset, Selgenes, order_list = order_list, palet = palet)
```

### Niche formation (Figure 3I) {.tabset}
#### Heatmap stroma
```{r heatmap LTBR TNFRSF14 NOTCH3 JAG1}
Selgenes <- c("LTBR","TNFRSF14","NOTCH3","JAG1")
order_list <-c("CAF1/PRC","SMC/PC", "CAF2/TRC","AdvFB")
data_subset <- subset(merged_data, cell_type  %in% order_list)
  
Heatmap(data_subset, Selgenes, order_list = order_list, palet = palet)
```

#### Heatmap CD8`r knitr::asis_output("\U207A")` T cell
```{r heatmap LTA LTB TNFSF14 DLL1 JAG1}
Selgenes <- c("LTA","LTB","TNFSF14","DLL1","JAG1")
order_list <-c(paste0("Stem-like/Naive CD8", "\u207A ", "T cells"), 
               paste0("Effector-memory CD8", "\u207A ", "T cells"),
               paste0("Effector CD8", "\u207A ", "T cells"),
               paste0("Exhausted CD8", "\u207A ", "T cells"),
               paste0("Cycling CD8", "\u207A ", "T cells") )
data_subset <- subset(merged_data, cell_type  %in% order_list)

Heatmap(data_subset, Selgenes, order_list = order_list, palet = palet)
```

### Heatmaps (Figure 3J) {.tabset}
#### LTBR regulated
```{r heatmap CLU TNFSF13B MFGE8}
Selgenes <- c("CLU","TNFSF13B","MFGE8")
order_list <-c("CAF1/PRC","SMC/PC", "CAF2/TRC","AdvFB")
data_subset <- subset(merged_data, cell_type  %in% order_list)
  
Heatmap(data_subset, Selgenes, order_list = order_list, palet = palet)
```

#### TNFRSF14 regulated
```{r heatmap DKK1 IL32 IL34}
Selgenes <- c("DKK1", "IL32", "IL34")
order_list <-c("CAF1/PRC","SMC/PC", "CAF2/TRC","AdvFB")
data_subset <- subset(merged_data, cell_type  %in% order_list)
  
Heatmap(data_subset, Selgenes, order_list = order_list, palet = palet)
```

#### NOTCH3 regulated
```{r heatmap HEY1 HEY2 HEYL}
Selgenes <- c("HEY1", "HEY2", "HEYL")
order_list <-c("CAF1/PRC","SMC/PC", "CAF2/TRC","AdvFB")
data_subset <- subset(merged_data, cell_type  %in% order_list)
  
Heatmap(data_subset, Selgenes, order_list = order_list, palet = palet)
```

### Session info
```{r session info}
sessionInfo()
date()
```
