---
title: "CCL19-expressing FRC subsets in coronavirus-vector based immunotherapy"
author: "Chrysa Papadopoulou"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: yes
    collapsed: no
    smooth_scroll: yes
mainfont: {Arial Unicode MS}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("ragg_png", "cairo_pdf"))
knitr::asis_output("\U207A")
knitr::spin_child('extra_functions.R')
set.seed(1000)
```

### Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(here)
  library(purrr)
  library(dplyr)
  library(stringr)
  library(patchwork)
  library(Seurat)
  library(Matrix)
  library(dittoSeq)
  library(gridExtra)
  library(gsubfn)
  library(ggsci)
  library(slingshot)
  library(tradeSeq)
  library(mgcv)
})
```

## Set directory
```{r set directory}
basedir <- here()
```

### Read CCL19-EYFP`r knitr::asis_output("\U207A")` mCOV-FIt31-g33 cell data
```{r read CCL19 FRC mCOV-FIt31-g33 data}
CCL19_EYFP_mCOV <- readRDS(paste0(basedir,"/data/Mouse/mCOV.rds"))
```

### Define color palette
```{r color vec}
cols <- c("#C77CFF","#F8766D","#00BA38","#B79F00","#FF64B0","#00BFC4","#00B4F0","#7CAE00")
names(cols) <-c(paste0("Smoc1", expression("\u207A "), "AdvFB"),paste0("Cd34", expression("\u207A "), "AdvFB"),paste0("Npnt", expression("\u207A "), "AlvFB"),paste0("Hhip", expression("\u207A "), "AdvFB"),paste0("Sulf1", expression("\u207A "), "TRC"),paste0("Rgs5", expression("\u207A "), "PRC"),"SMC/PC","TLS TRC")
```


### CCL19-EYFP`r knitr::asis_output("\U207A")` mCOV-FIt31-g33 cells
#### Umap colored per celltype (Supplementary Figure 5C)
```{r CCL19 EYFP umap mCOV-FIt31-g33 cells}
DimPlot(CCL19_EYFP_mCOV, reduction = "umap", group.by = "annot", cols=cols)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle(paste0("Ccl19-EYFP", "\U207A ", "cells (mCOV-FIt31-g33 cells)"))
```

### Dotplot (Figure 6F) {.tabset}
```{r dotplot mouse tumor mCOV scale true}
data_conv <-CCL19_EYFP_mCOV
data_conv <-Remove_ensebl_id(data_conv)

Idents(data_conv) <- data_conv$annot
levels(data_conv)<-levels(data_conv)[order(match(levels(data_conv),c(paste0("Cd34", expression("\u207A "), "AdvFB"),paste0("Smoc1", expression("\u207A "), "AdvFB"),paste0("Hhip", expression("\u207A "), "AdvFB"),paste0("Npnt", expression("\u207A "), "AlvFB"),paste0("Sulf1", expression("\u207A "), "TRC"),paste0("Rgs5", expression("\u207A "), "PRC"),"SMC/PC","TLS TRC")))]
data_conv$annot <- factor(as.character(data_conv@active.ident), levels = rev(c(paste0("Cd34", expression("\u207A "), "AdvFB"),paste0("Smoc1", expression("\u207A "), "AdvFB"), paste0("Hhip", expression("\u207A "), "AdvFB"),paste0("Npnt", expression("\u207A "), "AlvFB"),paste0("Sulf1", expression("\u207A "), "TRC"),paste0("Rgs5", expression("\u207A "), "PRC"),"SMC/PC","TLS TRC")))

gene_list <-c("Cd34","Pdpn","Hhip","Sulf1","Ccl11","Cxcl9","Tnc","Pdgfrb","Notch3","Rgs5","Acta2","Des","Myh11","Il33","Ccl19","Ccl21a","Cxcl13","Clu")

gg <-DotPlot(data_conv, features = gene_list, group.by = "annot", scale = TRUE, cols = c("lightgrey", "#C51B7D"), 
             scale.min = 0, scale.max = 100,col.min = 0, col.max = 8 , dot.scale = 4) + xlab(" ") + ylab(" ")

gg + theme(axis.text.x = element_text(angle = 45,hjust = 1))
```

### Differentiation trajectories on CCL19-EYFP`r knitr::asis_output("\U207A")` mCOV-FIt31-g33 cells based on Slingshot algorithm (Figure 6G) {.tabset}
#### Annotation from RNA_snn_res.0.25 to perform trajectory analysis
```{r umap CCL19 EYFP umap map annotation  mCOV-FIt31-g33}
#Set color palette
palet_new <- c("#C77CFF","#F8766D","#00BA38","#B79F00","#FF64B0","#00BFC4","#00B4F0","#7CAE00")
names(palet_new) <-c(2,1,4,3,5,6,7,8)

CCL19_EYFP_mCOV$annot_numeric <- CCL19_EYFP_mCOV$annot
CCL19_EYFP_mCOV$annot_numeric[which(CCL19_EYFP_mCOV$annot == paste0("Cd34", expression("\u207A "), "AdvFB"))] <- 1
CCL19_EYFP_mCOV$annot_numeric[which(CCL19_EYFP_mCOV$annot == paste0("Smoc1", expression("\u207A "), "AdvFB"))] <- 2
CCL19_EYFP_mCOV$annot_numeric[which(CCL19_EYFP_mCOV$annot == paste0("Hhip", expression("\u207A "), "AdvFB"))] <- 3
CCL19_EYFP_mCOV$annot_numeric[which(CCL19_EYFP_mCOV$annot == paste0("Npnt", expression("\u207A "), "AlvFB"))] <- 4
CCL19_EYFP_mCOV$annot_numeric[which(CCL19_EYFP_mCOV$annot == paste0("Sulf1", expression("\u207A "), "TRC"))] <- 5
CCL19_EYFP_mCOV$annot_numeric[which(CCL19_EYFP_mCOV$annot == paste0("Rgs5", expression("\u207A "), "PRC"))] <- 6
CCL19_EYFP_mCOV$annot_numeric[which(CCL19_EYFP_mCOV$annot == "SMC/PC")] <- 7
CCL19_EYFP_mCOV$annot_numeric[which(CCL19_EYFP_mCOV$annot == "TLS TRC")] <- 8

DimPlot(CCL19_EYFP_mCOV, reduction = "umap", group.by = "annot_numeric", cols = palet_new)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle("New annotation")

```

#### Differentiation trajectories for Sulf1`r knitr::asis_output("\U207A")` TRC, TLS TRC and Rgs5`r knitr::asis_output("\U207A")` PRC
```{r CCL19 EYFP mCOV-FIt31-g33 differentiation trajectories}
#Calculation of CCL19-EYFP`r knitr::asis_output("\U207A")` mCOV-FIt31-g33 cell differentiation trajectories
clustering <- CCL19_EYFP_mCOV@meta.data$annot_numeric
dimred <- CCL19_EYFP_mCOV@reductions$umap@cell.embeddings

# Slingshot for TRC (Sulf1, TLS)
pto_TRC <- slingshot(dimred, clustering, start.clus = '2', end.clus = '8', reducedDim = 'umap',extend="n",stretch=0.07,thresh=0.05)
pto_TRC <- as.SlingshotDataSet(pto_TRC)

# Slingshot for PRC
pto_PRC <- slingshot(dimred, clustering, start.clus = '7', end.clus = '6' ,reducedDim = 'umap',extend="n",stretch=0.07,thresh=0.05)
pto_PRC <- as.SlingshotDataSet(pto_PRC)

plot(dimred, col = palet_new[clustering], asp = 1, pch = 16)
lines(pto_PRC@curves$Lineage3, lwd = 3, col = 'magenta')
lines(pto_TRC@curves$Lineage3, lwd = 3, col = 'gold')
lines(pto_TRC@curves$Lineage1, lwd = 3, col = 'black')
legend("bottomright", legend = c("T3"), col = c('gold'), lty = 1, lwd = 2.5, cex = 1.1)
legend("topright", legend = names(cols), col = cols, pch = 2, pt.cex = 1, cex = 0.8)
```

### Combine gene signatures in T3 differentiation trajectory (Figure 6H)
```{r Step 1}
counts <- CCL19_EYFP_mCOV@assays$RNA@counts

CCL19_EYFP_mCOV <- FindVariableFeatures(CCL19_EYFP_mCOV, selection.method = "vst", nfeatures = 1000)

new_counts <-counts[rownames(counts) %in% CCL19_EYFP_mCOV@assays$RNA@var.features,]
sds <- as.SlingshotDataSet(pto_TRC)
pseudotime <- slingPseudotime(pto_TRC, na = FALSE)
cellWeights <- slingCurveWeights(pto_TRC)

BPPARAM <- BiocParallel::bpparam()
BPPARAM$workers <- 4 # use 2 cores
control <- gam.control()
control$maxit <- 1000 
```
```{r fitGAm, eval=FALSE}
set.seed(3)
sce <- fitGAM(counts = new_counts, pseudotime = pseudotime, cellWeights = cellWeights,
                 nknots = 6, verbose = TRUE,parallel=TRUE, BPPARAM = BPPARAM, sds = sds,control = control)

```

### Save sce object
```{r save sce}
#saveRDS(sce, paste0(basedir,"/data/Mouse/Trajectory_fitGAM_T3_1000.rds"))
```

### Read sce object
We provide the sce object because the fitGAM commnand took time to run
```{r read sce}
sce <- readRDS(paste0(basedir,"/data/Mouse/Trajectory_fitGAM_T3_1000.rds"))
```

We focus on genes whose expression changes across the T3 trajectory in the pseudotime. In particular, on those genes that are upregulated in TLS TRC. We use the startVsEndTest statistical test.

### Gene expression changes in T3 trajectory
```{r startVsEndTest}
pseudotime_start_end_association <- startVsEndTest(sce) 
```

### Calculate log-transformed counts and fitted values for a particular gene across the T3 lineage 
```{r Step 2}
data_Ccl21a <-plotSmoothers1(sce, counts, gene = "ENSMUSG00000094686.Ccl21a", lineagesToPlot = c(3))
data_Srgn <-plotSmoothers1(sce, counts, gene = "ENSMUSG00000020077.Srgn", lineagesToPlot = c(3)) 
data_Il33 <-plotSmoothers1(sce, counts, gene = "ENSMUSG00000024810.Il33", lineagesToPlot = c(3)) 
data_Cxcl12 <-plotSmoothers1(sce, counts, gene = "ENSMUSG00000061353.Cxcl12", lineagesToPlot = c(3)) 
data_Ccl19 <-plotSmoothers1(sce, counts, gene = "ENSMUSG00000071005.Ccl19", lineagesToPlot = c(3))

data_Ccl21a$gene <- "Ccl21a"
data_Srgn$gene <-"Srgn"
data_Il33$gene <-"Il33"
data_Cxcl12$gene <-"Cxcl12"
data_Ccl19$gene <-"Ccl19"
```

### Plot gene curves
```{r Step 3 plot}
cols_pal <- c("orange","#00B0F6","#00BB4E","#C09B00", "#FF6A98")
names(cols_pal) <-c("Ccl19","Il33","Cxcl12","Ccl21a","Srgn")

visual12 <- rbind(data_Ccl21a,data_Srgn,data_Il33,data_Cxcl12,data_Ccl19)
end_val <-round(max(visual12$time)) + 1
end_y_axis <-round(max(log1p(visual12$gene_count)))

ggplot(visual12, aes(x=time, y=log1p(gene_count), group=gene, col = gene, fill=gene)) +
  labs(x = "Pseudotime (T3)", y = "Log(expression + 1)") +
      geom_line(lwd = 2) +
  scale_y_continuous(expand = expansion(c(0,0)), limits = c(0.0, end_y_axis),breaks = c(0,1,2,3,4,end_y_axis)) +
  scale_x_continuous(limits = c(0.0, end_val + 0.5), expand = expansion(c(0,0)), breaks = c(0,5,end_val)) + 
 theme(aspect.ratio=1.3, axis.text.y = element_text(angle = 0,colour = "black")) + 
  theme(axis.line = element_line(colour = "black"),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
panel.border = element_blank(),
panel.background = element_blank(),
legend.title = element_blank(),axis.text.x = element_text(angle = 0, vjust = 0.5,colour = "black")) +
  scale_color_manual(values = cols_pal)+ 
  ggtitle(paste0("Cd34", "\U207A ", "AdvFB >> TLS TRC (T3)"))
```

### Multidimensional scaling (MDS) plot 

Multidimensional scaling (MDS) visualizes the level of similarity of variables in a data set. MDS recognizes the structure of the dataset in 2D, as it maintains the pairwise distances between data points.

### Read CCL19-EYFP`r knitr::asis_output("\U207A")` cell data from naïve lungs and excised LLC-gp33 tumors on day 23
```{r read CCL19 FRC data}
CCL19_EYFP <- readRDS(paste0(basedir,"/data/Mouse/CCL19_EYFP_nonmCOV.rds"))
```

### Merge CCL19-EYFP`r knitr::asis_output("\U207A")` cells from mCOV-FIt31-g33 and non immunization
```{r merge nomCOV mCOV data}
data_merge <- merge(CCL19_EYFP, y = c(CCL19_EYFP_mCOV),
             add.cell.ids = c("CCL19_EYFP","CCL19_EYFP_mCOV"),
             project = "merge_no_mcov_mcov")

#Preprocessing
resolution <- c(0.1, 0.25, 0.4, 0.6,0.8, 1.)
data_merge  <- preprocessing(data_merge,resolution)
```

### **Integrate data to correct for batch effects due to different condition (naive vs mCOV-FIt31-g33) via seurat**
### Step 1
```{r integrate data step 1}
#Split object by Patient as we see batch effects coming from different patients
obj.list <-SplitObject(data_merge, split.by = 'condition')
#For each object in list we see to run normalization and identify highly variable features
for (i in 1:length(obj.list)){
  #Normalization
  obj.list[[i]] <- NormalizeData(obj.list[[i]], normalization.method = "LogNormalize", scale.factor = 10000) 
  #Find high variable genes
  obj.list[[i]] <- FindVariableFeatures(obj.list[[i]], selection.method = "vst", nfeatures = 2000)#The purpose of this is to id
}
```

### Step 2
```{r integrate data step 2}
#select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = obj.list)
#Find anchors to integrate the data across different patients (Canonical correlation analysis)
anchors <- FindIntegrationAnchors(object.list = obj.list, anchor.features = features)
# this command creates an 'integrated' data assay
seurat_integrated <- IntegrateData(anchorset = anchors)
```

### Step 3
```{r integrate data step 3}
# We run a single integrated analysis on all cells!
DefaultAssay(seurat_integrated) <- "integrated"

# Run the standard workflow for visualization and clustering
seurat_integrated <- ScaleData(seurat_integrated, verbose = FALSE)
seurat_integrated <- RunPCA(object = seurat_integrated, npcs = 30, verbose = FALSE,seed.use = 8734)
seurat_integrated <- RunTSNE(object = seurat_integrated, reduction = "pca", dims = 1:20, seed.use = 8734)
seurat_integrated<- RunUMAP(object = seurat_integrated, reduction = "pca", dims = 1:20, seed.use = 8734)
 seurat_integrated <- FindNeighbors(object = seurat_integrated, reduction = "pca", dims = 1:20, seed.use = 8734)
#Clustering 
resolution <- c(0.1, 0.25, 0.4, 0.6,0.8, 1.)
for(k in 1:length(resolution)){
  seurat_integrated <- FindClusters(object = seurat_integrated, resolution = resolution[k], random.seed = 8734)
}
```

### Batch effects based on condition are corrected {.tabset}
#### Color per condition
```{r umap per condition}
DimPlot(seurat_integrated, reduction = "umap", group.by = "condition")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") 

```

#### Existing cell type annotation
```{r umap visualization}
DimPlot(seurat_integrated, reduction = "umap", group.by = "annot")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") 

```

### Plot gene signatures {.tabset}
```{r get full gene names}
DefaultAssay(seurat_integrated)<- "RNA"

genes <- c("NPNT","HHIP","CD34","ACTA2","SULF1","CLU","CCL21A","CCL19","IL33","SMOC1","CXCL13","RGS5")
genes <-str_to_sentence(genes)

genes <- unlist(lapply(genes, function(x) {
  get_full_gene_name(x,seurat_integrated)
}))
```

#### NPNT HHIP CD34 ACTA2
```{r NPNT HHIP CD34 ACTA2}
DefaultAssay(seurat_integrated)<- "RNA"

FeaturePlot(seurat_integrated, reduction = "umap", 
          features = genes[1:4],
          cols=c("lightgrey", "darkred"),
          order = T )+
theme(legend.position="right", legend.title=element_text(size=5)) 
```

#### SULF1 CLU CCL21A CCL19
```{r SULF1 CLU CCL21A CCL19}
DefaultAssay(seurat_integrated)<- "RNA"

FeaturePlot(seurat_integrated, reduction = "umap", 
          features = genes[5:8],
          cols=c("lightgrey", "darkred"),
          order = T )+
theme(legend.position="right", legend.title=element_text(size=5)) 
```

#### IL33 SMOC1 CXCL13 RGS5
```{r IL33 SMOC1 CXCL13 RGS5}
DefaultAssay(seurat_integrated)<- "RNA"

FeaturePlot(seurat_integrated, reduction = "umap", 
          features = genes[9:12],
          cols=c("lightgrey", "darkred"),
          order = T )+
theme(legend.position="right", legend.title=element_text(size=5)) 

```

### Perform new cell type annotation based on the previous gene signatures
```{r new cell type assignment}
DefaultAssay(seurat_integrated) <- "integrated"

seurat_integrated@meta.data$new_annot <--1
seurat_integrated$new_annot[which(seurat_integrated$integrated_snn_res.0.25 == 0)] <- paste0("Cd34", "\u207A ", "AdvFB")

seurat_integrated$new_annot[which(seurat_integrated$integrated_snn_res.0.25 == 1)] <- paste0("Hhip", "\u207A ", "AdvFB")

seurat_integrated$new_annot[which(seurat_integrated$integrated_snn_res.0.25 == 2)] <- paste0("Npnt", "\u207A ", "AlvFB")

seurat_integrated$new_annot[which(seurat_integrated$integrated_snn_res.0.25 == 3)] <- "SMC/PC"

seurat_integrated$new_annot[which(seurat_integrated$integrated_snn_res.0.25 == 4)] <- paste0("Sulf1", "\u207A ", "TRC")

seurat_integrated$new_annot[which(seurat_integrated$integrated_snn_res.0.25 == 5)] <- paste0("Rgs5", "\u207A ", "PRC")

seurat_integrated$new_annot[which(seurat_integrated$integrated_snn_res.0.25 == 6)] <- paste0("Npnt", "\u207A ", "AlvFB")

seurat_integrated$new_annot[which(seurat_integrated$integrated_snn_res.0.25 == 7)] <- paste0("TLS ", "TRC")
```

### Visualize new cell type annotation
```{r plot new cell type annotation}
DimPlot(seurat_integrated, reduction = "umap", group.by = "new_annot", cols = cols)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") #+ NoLegend()

```

### MDS computation
```{r MDS computation, , eval=FALSE}
# Before running MDS, we first calculate a distance matrix between all pairs of cells.
# Here we use a simple euclidean distance metric on all genes, using scale.data as input
d <- dist(t(GetAssayData(seurat_integrated, slot = 'scale.data')))
d_mat <-as.matrix(d)
# Run the MDS procedure, k determines the number of dimensions
mds <- cmdscale(d = d, k = 2)
# cmdscale returns the cell embeddings, we first label the columns to ensure downstream consistency
colnames(mds) <- paste0("MDS_", 1:2)
# We will now store this as a custom dimensional reduction called "mds"
seurat_integrated[['mds']] <- CreateDimReducObject(embeddings = mds, key = 'MDS_', assay = DefaultAssay(seurat_integrated))
```

```{r save integrated MDS data, eval=FALSE}
#saveRDS(seurat_integrated, paste0(basedir,"/data/Mouse/integrated_naive_mcov_mds.rds"))
```

We provide the integrated object with MDS representation as the classical MDS algorithm takes a long time to run.

### Read integrated object with MDS representation
```{r read integrated MDS data}
seurat_integrated <- readRDS(paste0(basedir,"/data/Mouse/integrated_naive_mcov_mds.rds"))
```

### Visualize MDS plot (Figure 6E)
```{r MDS visualization}
mds_tx_condition = seurat_integrated@reductions$mds@cell.embeddings %>% 
as.data.frame() %>% cbind(tx = seurat_integrated@meta.data$condition)

mds_tx_celltype = seurat_integrated@reductions$mds@cell.embeddings %>% 
as.data.frame() %>% cbind(tx = seurat_integrated@meta.data$new_annot)

mds_tx_TOTAL <-merge(mds_tx_condition, mds_tx_celltype, by=c("MDS_1", "MDS_2"), all.x=T, all.y=T)
colnames(mds_tx_TOTAL) <-c("MDS_1", "MDS_2", "Condition","Celltype")

mds_tx_TOTAL <- mds_tx_TOTAL %>% 
  group_by(Celltype,Condition) %>% 
  mutate(count_mds1 = mean(MDS_1)) %>%
  mutate(count_mds2 = mean(MDS_2))

test <-mds_tx_TOTAL %>%
  group_by(Celltype) %>%
  filter(!(Celltype == 'TLS TRC' & Condition == 'Not immunized'))

ggplot(test, aes(x=count_mds1, y=count_mds2, color=Celltype, shape = Condition)) + geom_point(stroke = 1.5) + ylab("MDS2") + xlab("MDS1") +  
scale_color_manual(values=cols) + scale_shape_manual(values = c(2, 3)) + 
  theme(aspect.ratio = 1.3, axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text.y = element_text(angle = 0, vjust = 0.5,colour = "black",size = 10),
    axis.text.x = element_text(angle = 0, vjust = 0.5,colour = "black",size = 10),
    
) 
```

### Session info
```{r session info}
sessionInfo()
date()
```
