---
title: "NSCLC stroma and CCL19 cells"
author: "Chrysa Papadopoulou"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: yes
    collapsed: no
    smooth_scroll: yes
mainfont: {Arial Unicode MS}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("ragg_png", "cairo_pdf"))
knitr::asis_output("\U207A")
knitr::spin_child('extra_functions.R')
set.seed(1000)
```

### Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(here)
  library(purrr)
  library(dplyr)
  library(stringr)
  library(patchwork)
  library(Seurat)
  library(Matrix)
  library(dittoSeq)
  library(gridExtra)
  library(gsubfn)
  library(ggsci)
})
```

## **Stroma cells in NSCLC**

## Set directory
```{r set directory}
basedir <- here()
```

### Read Stroma cell data
```{r read stroma cell data}
data <- readRDS(paste0(basedir,"/data/Human/NSCLC_stroma_total.rds"))
```

### Define color palette
```{r color vector}
cols<- pal_igv()(51)
names(cols) <- c(0:50)
```

### NSCLC Stroma cells (Supplementary Figure 2A-2B){.tabset}
#### Patients

```{r umap per patient}
# Total fibroblasts and endothelial cells across NSCLC patients
colors_pID <-c("#F8766D","#00C08B","#00B4F0","#0ADD08","#B79F00")
names(colors_pID) <-c("NSCLC#2","NSCLC#3","NSCLC#4","NSCLC#6","NSCLC#7")

DimPlot(data, reduction = "tsne", group.by = "patient", cols=colors_pID)+ 
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlab("tSNE1") +
  ylab("tSNE2") + ggtitle("Patients")

```

#### Origin
```{r umap per origin}
# Total fibroblasts and endothelial cells isolated from SM, CM and unaffected lung (LU)
colors_origin <-c("red","blue" ,"#33CC00FF") 
names(colors_origin) <- c("Subpleural Margin","Lung (unaffected)","Central Margin")

DimPlot(data, reduction = "tsne", group.by = "origin", cols=colors_origin )+ 
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  xlab("tSNE1") +
  ylab("tSNE2") + ggtitle("Origin")
```

#### COL1A2
```{r feature plot COL1A2}
FeaturePlot(data, reduction = "tsne", 
          features = get_full_gene_name('COL1A2',data),raster=FALSE,
          cols=c("lightgrey", "darkred")) + ggtitle("CAF/FB(COL1A2)")
```

#### PECAM1
```{r feature plot PECAM1}
FeaturePlot(data, reduction = "tsne", 
          features = get_full_gene_name('PECAM1',data),raster=FALSE,
          cols=c("lightgrey", "darkred")) + ggtitle("EC(PECAM1)")
```

#### Cell type
```{r umap per cell type}
palet <- cols[4:10]
names(palet) <- c("CAF2","CAF1", "EC", "FB" ,"Meso","SMC/PC")

DimPlot(data, reduction = "tsne", group.by = "cell_type", cols= palet)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
         panel.grid.major = element_blank()) +
 xlab("TSNE1") +
ylab("TSNE2") 
```

### Dotplots (Supplementary Figure 2C and 2F) {.tabset}
#### Cell type assignment 
```{r dotplot stroma}
data_conv <-data
data_conv <-Remove_ensebl_id(data_conv)

Idents(data_conv) <- data_conv$cell_type
levels(data_conv)<-levels(data_conv)[order(match(levels(data_conv),c("CAF2","FB","CAF1","SMC/PC","Meso","EC")))]
data_conv$cell_type <- factor(as.character(data_conv@active.ident), levels = rev(c("CAF2","FB","CAF1","SMC/PC","Meso","EC")))

gene_list <-c("COL1A2","POSTN","MMP2","PDPN","PDGFRA","PDGFRB","ACTA2","RGS5","KRT19","PECAM1")

gg <- dittoDotPlot(data_conv, vars = gene_list, group.by = "cell_type", size = 9,legend.size.title = "% expressed",scale = FALSE,summary.fxn.color = mean, max = 6.5, min = 0 , min.color = "#D1E5F0" , max.color = "#631879FF")

gg + ggtitle("Celltype assignment")
```

#### Chemokines
```{r dotplot chemokines}
gene_list <-c("CCL19","CCL21","CCL3","CCL5","CCL8","CXCL10","CXCL3","CXCL9")

gg <- dittoDotPlot(data_conv, vars = gene_list, group.by = "cell_type", size = 9,legend.size.title = "% expressed",scale = FALSE,summary.fxn.color = mean, min.percent = 0.02, max.percent = 0.7, max = 1.6, min = 0 ,min.color = "#D1E5F0" , max.color = "#631879FF")

gg + ggtitle("Chemokines")
```

### Stacked Barplots (Supplementary Figure 2D and 2E) {.tabset}
#### Number of cells per patient 
```{r stacked barplot cells per patient}
df <-data@meta.data %>% count(patient, cell_type) %>%    # Group by patient and cell_type, then count number in each group
         mutate(pct=n/sum(n))                            # Calculate percent within each patient

df$cell_type <- factor(df$cell_type, levels=names(palet))
df$patient <- factor(df$patient, levels=c("NSCLC#2", "NSCLC#3" ,"NSCLC#4" ,"NSCLC#6" ,"NSCLC#7"))

ggplot(df, aes(patient, n, fill=cell_type)) +
  geom_bar(stat="identity") +
  theme(  axis.line = element_line(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()) +
  scale_y_continuous(expand = c(0, 0))+
  labs(y= "Cells", x= " ") +
  scale_fill_manual(values = palet)

```

#### Number of cells per origin
```{r stacked barplot cells per origin}
df <-data@meta.data %>% count(origin, cell_type) %>%    # Group by orign and cell_type, then count number in each group
         mutate(pct=n/sum(n))                           # Calculate percent within each patient

df$cell_type <- factor(df$cell_type, levels=names(palet))
df$origin <- factor(df$origin, levels=c("Lung (unaffected)", "Subpleural Margin", "Central Margin"))

ggplot(df,aes(origin, n, fill=cell_type)) +
  geom_bar(stat="identity") +
  theme(  axis.line = element_line(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()) +
  scale_y_continuous(expand = c(0, 0))+
  labs(y= "Cells", x= " ") +
  scale_fill_manual(values = palet)
```

```{r deallocate memory NSCLC stroma data, include=FALSE}
remove(data)
```

## **CCL19`r knitr::asis_output("\U207A")` FRCs in NSCLC**

### Read CCL19`r knitr::asis_output("\U207A")` FRC data
```{r read CCL19 FRC data}
NSCLC_CCL19_data <- readRDS(paste0(basedir,"/data/Human/NSCLC_CCL19_FRCs_CAFs.rds"))
```

### CCL19`r knitr::asis_output("\U207A")` fibroblasts (Figure 1K)
```{r umap CCL19 FRCs)}
#Define color palet
palet_CCL19_FRC <- c("#1B9E77", "#54B0E4","#E3BE00", "#E41A1C")
names(palet_CCL19_FRC) <- c("CAF2/TRC","CAF1/PRC","AdvFB" ,"SMC/PC")

palet_CCL19_FRC <- palet_CCL19_FRC[names(palet_CCL19_FRC) %in% unique(NSCLC_CCL19_data$cell_type)]

DimPlot(NSCLC_CCL19_data, reduction = "umap", group.by = "cell_type",cols = palet_CCL19_FRC)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle(paste0("CCL19", "\U207A ", "fibroblasts"))

```

### Dotplot (Figure 1L)
```{r dotplot}
data_conv <-NSCLC_CCL19_data

data_conv <-Remove_ensebl_id(data_conv)

Idents(data_conv) <- data_conv$cell_type
levels(data_conv)<-levels(data_conv)[order(match(levels(data_conv),c("SMC/PC","CAF1/PRC","AdvFB","CAF2/TRC")))]
data_conv$cell_type <- factor(as.character(data_conv@active.ident), levels = rev(c("SMC/PC","CAF1/PRC","AdvFB","CAF2/TRC")))

gene_list <-c("CCL19","CCL21","PDPN","FAP","POSTN","CLU","LEPR","CD34","SULF1","DPT","ICAM1","VCAM1","ACTA2","MYH11",
"MCAM","NOTCH3","RGS5","DES","AIFM2")

dittoDotPlot(data_conv, vars = gene_list, group.by = "cell_type", size = 8,legend.size.title = "Expression (%)",scale = FALSE) + ylab(" ") + ggtitle(paste0("CCL19", expression("\u207A"), " fibroblasts"))

```

### Signatures {.tabset}
#### SLO-PRC (Figure 1M)
```{r sign SLO-PRC}
SLO_PRC <-list("CCL19","CCL21","ITGA1","ITGA7","MCAM","CNN1","NOTCH3","ACTA2","PDGFRB","ANGPT2")
object <- AddModuleScore(object = data_conv, features = SLO_PRC, name = "SLO_PRC_signature",ctrl = 20)
FeaturePlot(object = object, features = "SLO_PRC_signature10",min.cutoff = -1, max.cutoff = 2.5) + ggtitle("SLO-PRC signature")
```

#### SLO-TRC (Figure 1N)
```{r sign SLO-TRC}
SLO_TRC <-c("CCL19","CCL21","PDPN","ICAM1","VCAM1","LUM","PDGFRA","TNFSF13B")
object <- AddModuleScore(object = data_conv, features = SLO_TRC, name = "SLO_TRC_signature",ctrl = 20)
FeaturePlot(object = object, features = "SLO_TRC_signature8",min.cutoff = -1, max.cutoff = 2.5) + ggtitle("SLO-TRC signature")
```

### Signatures (Supplementary Figure 2G) {.tabset}
#### CCL21
```{r feature plot CCL21}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('CCL21',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("CCL21")
```

#### ACTA2
```{r feature plot ACTA2}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('ACTA2',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("ACTA2")
```

#### ITGA1
```{r feature plot ITGA1}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('ITGA1',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("ITGA1")
```

#### NOTCH3
```{r feature plot NOTCH3}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('NOTCH3',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("NOTCH3")
```

#### MCAM
```{r feature plot MCAM}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('MCAM',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("MCAM")
```

#### CCL19
```{r feature plot CCL19}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('CCL19',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("CCL19")
```

#### PDPN
```{r feature plot PDPN}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('PDPN',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("PDPN")
```

#### ICAM1
```{r feature plot ICAM1}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('ICAM1',NSCLC_CCL19_data)[2],raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("ICAM1")
```

#### VCAM1
```{r feature plot VCAM1}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('VCAM1',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("VCAM1")
```

#### LUM
```{r feature plot LUM}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('LUM',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("LUM")
```

### Extract NSCLC TRC and PRC
```{r subset NSCLC TRC PRC}
NCLS_FRCS <- subset(NSCLC_CCL19_data, cell_type %in% c("CAF2/TRC","CAF1/PRC"))

#Preprocessing
resolution <- c(0.1, 0.25, 0.4, 0.6,0.7, 0.8, 0.9, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0)
NCLS_FRCS <- FindVariableFeatures(NCLS_FRCS, selection.method = "vst", nfeatures = 2000)
NCLS_FRCS <- ScaleData(NCLS_FRCS)
NCLS_FRCS <- RunPCA(object = NCLS_FRCS, npcs = 30, verbose = FALSE,seed.use = 8734)
NCLS_FRCS <- RunTSNE(object = NCLS_FRCS, reduction = "pca", dims = 1:20, seed.use = 8734)
NCLS_FRCS <- RunUMAP(object = NCLS_FRCS, reduction = "pca", dims = 1:20, seed.use = 8734)
NCLS_FRCS <- FindNeighbors(object = NCLS_FRCS, reduction = "pca", dims = 1:20, seed.use = 8734)
for(k in 1:length(resolution)){
  NCLS_FRCS <- FindClusters(object = NCLS_FRCS, resolution = resolution[k], random.seed = 8734)
}

```

```{r deallocate memory CCL19 data, include=FALSE}
remove(NSCLC_CCL19_data)
```

### CCL19`r knitr::asis_output("\U207A")` TRC and PRC
```{r umap CCL19 TRC PRC, include=FALSE}
#Define color palette
palet <- palet_CCL19_FRC[names(palet_CCL19_FRC) %in% unique(NCLS_FRCS$cell_type)]

DimPlot(NCLS_FRCS, reduction = "umap", group.by = "cell_type",cols = palet)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### Save NSCLC CCL19`r knitr::asis_output("\U207A")` TRC PRC data
```{r save CCL19 TRC PRC}
#saveRDS(NCLS_FRCS, paste0(basedir,"/data/Human/NSCLC_CCL19_TRC_PRC_CAFs.rds"))
```

### Session info
```{r session info}
sessionInfo()
date()
```
