---
title: "Differentiation anaylsis in NSCLC"
author: "Chrysa Papadopoulou, Almut Lütge"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: yes
    collapsed: no
    smooth_scroll: yes
mainfont: {Arial Unicode MS}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("ragg_png", "cairo_pdf"))
knitr::asis_output("\U207A")
knitr::spin_child('extra_functions.R')
set.seed(1000)
```

### Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(here)
  library(purrr)
  library(dplyr)
  library(stringr)
  library(patchwork)
  library(Seurat)
  library(Matrix)
  library(ggsci)
  library(monocle3)
  library(SeuratWrappers)
  library(ggplot2)
  library(scCustomize)
})
```

## **Trajectory analysis in NSCLC**

## Set directory
```{r set directory}
basedir <- here()
```

### Read CCL19`r knitr::asis_output("\U207A")` FRC data
```{r read CCL19 FRC data}
NSCLC_CCL19_data <- readRDS(paste0(basedir,"/data/Human/NSCLC_CCL19_FRCs_CAFs.rds"))
```

### Color palette per celltype
```{r color palet celltype}
palet_CCL19_FRC <- c("#1B9E77", "#54B0E4","#E3BE00", "#E41A1C")
names(palet_CCL19_FRC) <- c("CAF2/TRC","CAF1/PRC","AdvFB" ,"SMC/PC")
```

### CCL19`r knitr::asis_output("\U207A")` fibroblasts
```{r umap CCL19 FRCs)}
palet_region <- c("#C77CFF","#B79F00","#F8766D")
names(palet_region) <- c("Central Margin","Subpleural Margin","Lung (unaffected)")

DimPlot(NSCLC_CCL19_data, reduction = "umap", group.by = "group",cols = palet_region)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle(paste0("CCL19", "\U207A ", "fibroblasts"))

```


### **Differentiation trajectories on NSCLC CCL19`r knitr::asis_output("\U207A")` cells based on Monocle3**

### Monocle3 preparation
```{r monocle3 preparation}
sceSub <-as.SingleCellExperiment(NSCLC_CCL19_data)

sce_list <- list("all" = sceSub)
sce_names <- names(sce_list)

test = rownames(rowData(sce_list$all))
test_new = as.data.frame(test)

test_new <- gsub("^.*\\.", "", test)

cds_list <- lapply(sce_names, prep_sce) %>% set_names(sce_names)
```

### Visulaize partitions
```{r show partitions}
for (sce_nam in sce_names) {
  p <-plot_cells(cds_list[[sce_nam]], color_cells_by = "partition")
  print(p)
}
```

### Learn graph
```{r learn graph}
cds_list <- lapply(cds_list, function(cds){
  cds <- learn_graph(cds, learn_graph_control = list(minimal_branch_len = 30,
                                                   L1.gamma = 0.08,
                                                   L1.sigma = 0.01,
                                                   nn.n_trees = 20
                                                   )) 
  })
```

### Learn pseudotime
```{r learn pseudotime}
# Written by Almut Lütge
cds_list <- lapply(cds_list, function(cds){
  mu_start <- get_earliest_principal_node(cds, "SMC/PC")
  lu_start <- get_earliest_principal_node(cds, "AdvFB")
  cds <- order_cells(cds, root_pr_nodes=c(mu_start, lu_start))
}) %>% set_names(sce_names)

```

### Visualize trajectories {.tabset}
#### Origin (Figure 4A)
```{r show trajectories origin}
for (sce_nam in sce_names) {
  p <- plot_trajectory(col_var = "group", cds = cds_list[[sce_nam]], palet_region)
  print(p)
}
```

#### Visualize pseudotime (Figure 4B)
```{r show pseudotime}
for (sce_nam in sce_names) {
  p <- plot_pseudo(cds_list[[sce_nam]])
  print(p)
}
```

#### Celltype
```{r show trajectories celltype}
for (sce_nam in sce_names) {
  p <- plot_trajectory(col_var = "cell_type", cds = cds_list[[sce_nam]], palet_CCL19_FRC)
  print(p)
}
```

### Visualize genes along pseudotime {.tabset}
#### DES MYH11 (Figure 4C)
```{r DES MYH11}
for (sce_nam in sce_names) {
  p <- plot_top_de(sce_nam, "1", gene_list = c("DES", "MYH11"))
  print(p)
}
```

#### RGS5 NOTCH3 ACTA2 (Figure 4E)
```{r RGS5 NOTCH3 ACTA2}
for (sce_nam in sce_names) {
  p <- plot_top_de(sce_nam, "1", gene_list = c("RGS5","NOTCH3","ACTA2"))
  print(p)
}
```

#### TNC CCL21 CXCL12 (Figure 4G)
```{r TNC CCL21 CXCL12}
for (sce_nam in sce_names) {
  p <- plot_top_de(sce_nam, "1", gene_list = c("TNC","CCL21","CXCL12"))
  print(p)
}
```

#### CLU POSTN SULF1 (Figure 4I)
```{r CLU POSTN SULF1}
for (sce_nam in sce_names) {
  p <- plot_top_de(sce_nam, "1", gene_list = c("CLU","POSTN","SULF1"))
  print(p)
}
```

### Unaffected Lung vs Tumor (Supplementary Figure 7A and 7B) {.tabset}
#### Umap
```{r split per tissue type umap}
DimPlot(NSCLC_CCL19_data, reduction = "umap", group.by = "type", split.by = "type")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

#### DES
```{r split per tissue type DES}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", split.by = "type",
          features = get_full_gene_name('DES',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5)
```

#### ACTA2
```{r split per tissue type ACTA2}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", split.by = "type",
          features = get_full_gene_name('ACTA2',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5)
```


### Signatures (Supplementary Figure 7E) {.tabset}
#### CD34
```{r feature plot CD34}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('CD34',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("CD34")
```

#### LEPR
```{r feature plot LEPR}
FeaturePlot(NSCLC_CCL19_data, reduction = "umap", 
          features = get_full_gene_name('LEPR',NSCLC_CCL19_data),raster=FALSE,
          cols=c("lightgrey", "darkred"), min.cutoff = 0, max.cutoff = 4.5) + ggtitle("LEPR")
```


### Session info
```{r session info}
sessionInfo()
date()
```