---
title: "NSCLC and BRCA CAFs"
author: "Chrysa Papadopoulou"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
mainfont: {Arial Unicode MS}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("ragg_png", "cairo_pdf"))
knitr::asis_output("\U207A")
knitr::spin_child('extra_functions.R')
set.seed(1000)
```

### Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(here)
  library(purrr)
  library(dplyr)
  library(stringr)
  library(patchwork)
  library(Seurat)
  library(Matrix)
  library(dittoSeq)
  library(gridExtra)
  library(gsubfn)
  library(ggsci)
  library(bigmds)
  library(tidyverse)
})
```

## **Comparison between NSCLC CCL19`r knitr::asis_output("\U207A")` FRCs, NSCLC CCL19`r knitr::asis_output("\U207A")` CAFs and BRCA CAFs**

### NSCLC CCL19`r knitr::asis_output("\U207A")` FRCs and NSCLC CCL19`r knitr::asis_output("\U207A")` CAFs

### Set directory
```{r set directory}
basedir <- here()
```

### Read CCL19 FRC data
```{r read CCL19 FRC data}
NSCLC_CCL19_data <- readRDS(paste0(basedir,"/data/Human/NSCLC_CCL19_FRCs_CAFs.rds"))
```

### Read NSCLC CCL19`r knitr::asis_output("\U207A")` TRC PRC data
```{r read  CCL19 TRC PRC}
NSCLC_CCL19_TRC_PRC <- readRDS(paste0(basedir,"/data/Human/NSCLC_CCL19_TRC_PRC_CAFs.rds"))
```

### Define color palette
```{r color palet}
palet <- c("#1B9E77", "#54B0E4","#E3BE00", "#E41A1C")
names(palet) <- c("CAF2/TRC","CAF1/PRC","AdvFB" ,"SMC/PC")
```

### Dotplot with apCAF, iCAF and myCAF gene signatures (Supplementary Figure 3C)
```{r dotplot, fig.width=17, fig.height=4}
data_conv <-NSCLC_CCL19_TRC_PRC
data_conv <-Remove_ensebl_id(data_conv)

CAF_subsets <- list("iCAF genes"= c("EFEMP1","IL6","C3","CFD","CLU","CXCL12","TNXB",
                                    "HAS1","PLA2G2A","GSN","PCOLCE2","CD34","LEPR",
                                    "CXCL14","CCL2","ADH1B","GPC3","VCAM1","TNC","ALDH1A2"),
                     "apCAF genes" = c("CD74","HLA-DRB1","HLA-DRA","HLA-DQB1"),
                    "myCAF genes"=c("FAP","POSTN","LRRC15","MMP11","COL10A1","COMP","COL8A1","GREM1","SULF1","COL13A1",
                    "COL5A2","LRRC17","COL12A1","THBS2","GJB2","ACTA2","MYH11","ACTG2","DES","COL4A6","BCAM",
                    "RGS5","MCAM","COL4A1","COL18A1","LAMC3","ARHGDIB"))

DotPlot(object = data_conv, features=CAF_subsets,group.by = "cell_type" ,scale=FALSE,dot.scale = 4) +  theme(strip.placement = "outside") + 
  theme(axis.text.x = element_text(angle = 90,hjust = 1,colour = NULL,face="bold"),axis.title.x.top=element_text(hjust = 1,face="bold")) +xlab(" ")+ylab(" ")
```

### CAF Signatures (Supplementary Figure 3D) {.tabset}
#### apCAF signature
```{r apCAF sign, results='hide'}
apCAF_genes <-c("CD74","HLA-DRB1","HLA-DRA","HLA-DQB1")

#Get full gene name (together with the ensembl id)
apCAF_genes <- unlist(lapply(apCAF_genes, function(x) {
  get_full_gene_name(x,NSCLC_CCL19_data)
})) 

slot_type <-"data"
gn <- "apCAF"
Visualize_GeneSignatures_sc(NSCLC_CCL19_data, apCAF_genes, slot_type, 'average.mean',gn) + ggtitle("apCAF signature")
```

#### iCAF signature 
```{r iCAF sign, results='hide'}
iCAF_genes <-c("EFEMP1","IL6","C3","CFD","CLU","CXCL12","TNXB","HAS1","PLA2G2A","GSN","PCOLCE2","CD34","LEPR","CXCL14","CCL2","ADH1B","GPC3","VCAM1","TNC","ALDH1A2")

#Get full gene name (together with the ensembl id)
iCAF_genes <- unlist(lapply(iCAF_genes, function(x) {
  get_full_gene_name(x,NSCLC_CCL19_data)
})) 

slot_type <-"data"
gn <- "iCAF"
Visualize_GeneSignatures_sc(NSCLC_CCL19_data, iCAF_genes, slot_type, 'average.mean',gn) + ggtitle("iCAF signature")
```


#### myCAF signature
```{r myCAF sign, results='hide'}
myCAF_genes <-c("FAP","POSTN","LRRC15","Mmp11","COL10A1","COMP","COL8A1","GREM1","SULF1","COL13A1","COL5A2","LRRC17","COL12A1","Thbs2","GJB2","ACTA2","MYH11","ACTG2","DES","COL4A6","BCAM","RGS5","MCAM","COL4A1","COL18A1","LAMC3","ARHGDIB")

#Get full gene name (together with the ensembl id)
myCAF_genes <- unlist(lapply(myCAF_genes, function(x) {
  get_full_gene_name(x,NSCLC_CCL19_data)
})) 

slot_type <-"data"
gn <- "myCAF"
Visualize_GeneSignatures_sc(NSCLC_CCL19_data, myCAF_genes, slot_type, 'average.mean',gn) + ggtitle("myCAF signature")
```


### BRCA (breast cancer) CAFs

### Read BRCA data from Cords et al, 2023
```{r read breast stroma cell data}
data_breast <- readRDS(paste0(basedir,"/data/Public/BREAST_fibro_tumour.rds"))
```

### Define color palette
```{r color vector}
cols<- pal_igv()(51)
names(cols) <- c(0:50)
```

### CAF annotation
```{r umap CAFs breast}
DimPlot(data_breast, reduction = "umap", group.by = "CAFtype")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle(paste0("CAFs in breast cancer"))
```

### Signatures related to SLOs (Supplementary Figure 3E)  {.tabset}
#### CCL19
```{r feature plot CCL19 total data breast}
FeaturePlot(data_breast, reduction = "umap", 
          features = "CCL19",
          cols=c("lightgrey", "darkred")) + ggtitle("CAF (BRCA)")
```

### Subset on CCL19`r knitr::asis_output("\U207A")` cells
```{r subset on CCL19 and preprocessing}
ccl19_data<-subset(x=data_breast,subset=CCL19>0, invert=FALSE)


resolution <- c(0.1, 0.25, 0.4, 0.5,0.6,0.8, 1.,1.2,1.4,1.6,1.8,2.)

# run sctransform
ccl19_data  <- SCTransform(ccl19_data, vars.to.regress = "percent.mt", verbose = FALSE)
ccl19_data <- RunPCA(object = ccl19_data, assay = "SCT",npcs = 30, verbose = FALSE,seed.use = 8734)
ccl19_data <- RunTSNE(object = ccl19_data, assay = "SCT",reduction = "pca", dims = 1:20, seed.use = 8734)
ccl19_data <- RunUMAP(object = ccl19_data, assay = "SCT", reduction = "pca", dims = 1:20, seed.use = 8734)
ccl19_data <- FindNeighbors(object = ccl19_data, reduction = "pca", dims = 1:20, seed.use = 8734)
for(k in 1:length(resolution)){
  ccl19_data <- FindClusters(object = ccl19_data, resolution = resolution[k], random.seed = 8734)
}
```

### Annotate CCL19`r knitr::asis_output("\U207A")` breast cancer data
```{r data annotation}
ccl19_data$cell_type <- -1
ccl19_data$cell_type[which(ccl19_data$SCT_snn_res.0.1 == 1)] <- "CAF1/PRC"
ccl19_data$cell_type[which(ccl19_data$SCT_snn_res.0.1 == 0)] <- "CAF2/TRC"
ccl19_data$cell_type[which(ccl19_data$SCT_snn_res.0.1 == 2)] <- "CAF2/TRC"
```

### CCL19`r knitr::asis_output("\U207A")` fibroblasts (BRCA) (Supplementary Figure 3F)
```{r umap CCL19 TRC PRC}
palet <- cols[4:10]
names(palet) <- c("CAF1/PRC","CAF2/TRC")

DimPlot(ccl19_data, reduction = "umap", group.by = "cell_type", cols= palet)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
         panel.grid.major = element_blank()) +
 xlab("UMAP1") +
  ylab("UMAP2") + ggtitle(paste0("CCL19", expression("\u207A"), " fibroblasts (BRCA)"))
```

### Dotplot (Supplementary Figure 3G)
```{r dotplot breast cancer}
data_conv <-ccl19_data

Idents(data_conv) <- data_conv$cell_type
levels(data_conv)<-levels(data_conv)[order(match(levels(data_conv),c("CAF1/PRC","CAF2/TRC")))]
data_conv$cell_type <- factor(as.character(data_conv@active.ident), levels = rev(c("CAF1/PRC","CAF2/TRC")))

gene_list <-c("CCL19","CCL21","PDPN","FAP","POSTN","CLU","LEPR","CD34","SULF1","DPT","ICAM1","VCAM1","ACTA2","MYH11","MCAM","NOTCH3","RGS5","DES")

dittoDotPlot(data_conv, vars = gene_list, group.by = "cell_type", size = 8,legend.size.title = "Expression (%)",scale = FALSE, max = 3.5) + ylab(" ")
```


### Session info
```{r session info}
sessionInfo()
date()
```