---
title: "NSCLC TILs"
author: "Chrysa Papadopoulou"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
mainfont: {Arial Unicode MS}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("ragg_png","cairo_pdf"))
knitr::asis_output("\U207A")
knitr::spin_child('extra_functions.R')
set.seed(1000)
```

## Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(here)
  library(purrr)
  library(stringr)
  library(patchwork)
  library(Seurat)
  library(Matrix)
  library(dittoSeq)
  library(gridExtra)
  library(dplyr)
})
```

## **Infiltrating lymphocytes in NSCLC**

### Set directory
```{r set directory}
basedir <- here()
```

### Read NSCLC TIL data
```{r read NSCLC TIL}
NSCLS_TIL_data <-readRDS(paste0(basedir,"/data/Human/NSCLC_TILs.rds"))
```

### NSCLC infiltrating lymphocytes (Supplementary Figure 3C) {.tabset}
#### umap 
```{r NSCLC TILS}
#Define color palet
palet <- c("#1B9E77", "#54B0E4","#E3BE00", "#E41A1C","#4DAF4A","#377EB8","#A65628","#222F75","#FB9A99")
names(palet) <- c("TRC","PRC","AdvFB" ,"SMC/PC",  paste0("CD4", "\u207A ", "T cells"),"B cells", "Regulatory T cells" ,paste0("Cycling CD8", "\u207A ", "T cells"),paste0("CD8", "\u207A ", "T cells"))

palet <- palet[names(palet) %in% unique(NSCLS_TIL_data$cell_type)]

DimPlot(NSCLS_TIL_data, reduction = "umap", group.by = "cell_type",cols = palet)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle("NSCLC infiltrating lymphocytes")

```

#### CD4
```{r CD4 expr, results='hide'}

gg <- FeaturePlot(NSCLS_TIL_data, reduction = "umap", 
          features =  get_full_gene_name("CD4",NSCLS_TIL_data)[4],
          cols=c("lightgrey", "darkred"),
          order = T,combin = FALSE)

 
gg[[1]]$labels$title <- "CD4"
gg
```

#### CD8A
```{r CD8A expr, results='hide'}
gg <-FeaturePlot(NSCLS_TIL_data, reduction = "umap", 
          features =  get_full_gene_name("CD8A",NSCLS_TIL_data),
          cols=c("lightgrey", "darkred"),
          order = T,combin = FALSE)

gg[[1]]$labels$title <- "CD8A"
gg
```

#### CD79A
```{r CD79A expr, results='hide'}
gg <-FeaturePlot(NSCLS_TIL_data, reduction = "umap", 
          features =  get_full_gene_name("CD79A",NSCLS_TIL_data),
          cols=c("lightgrey", "darkred"),
          order = T,combin = FALSE) 

gg[[1]]$labels$title <- "CD79A"
gg
```

### Read NSCLS CCL19`r knitr::asis_output("\U207A")` FRCs and NSCLS TILs
```{r read NSCLC FRC TIL}
merged_data <-readRDS(paste0(basedir,"/data/Human/NSCLC_TILs_SI3.rds"))
```

### **CD8`r knitr::asis_output("\U207A")` T cell subsets in NSCLC**
```{r subset CD8 T}
NCLS_TIL_FRC_cd8 <- subset(merged_data, cell_type %in% c(paste0("CD8", "\u207A ", "T cells"), paste0("Cycling CD8", "\u207A ", "T cells")))

#Preprocessing
resolution <- c(0.1, 0.25, 0.4, 0.6,0.7, 0.8, 0.9, 1.0, 1.2, 1.4, 1.6, 1.8, 2.0)
NCLS_TIL_FRC_cd8 <- FindVariableFeatures(NCLS_TIL_FRC_cd8, selection.method = "vst", nfeatures = 2000)
NCLS_TIL_FRC_cd8 <- ScaleData(NCLS_TIL_FRC_cd8)
NCLS_TIL_FRC_cd8 <- RunPCA(object = NCLS_TIL_FRC_cd8, npcs = 30, verbose = FALSE,seed.use = 8734)
NCLS_TIL_FRC_cd8 <- RunTSNE(object = NCLS_TIL_FRC_cd8, reduction = "pca", dims = 1:20, seed.use = 8734)
NCLS_TIL_FRC_cd8 <- RunUMAP(object = NCLS_TIL_FRC_cd8, reduction = "pca", dims = 1:20, seed.use = 8734)
NCLS_TIL_FRC_cd8 <- FindNeighbors(object = NCLS_TIL_FRC_cd8, reduction = "pca", dims = 1:20, seed.use = 8734)
for(k in 1:length(resolution)){
  NCLS_TIL_FRC_cd8 <- FindClusters(object = NCLS_TIL_FRC_cd8, resolution = resolution[k], random.seed = 8734)
}
```

### Annotation of CD8`r knitr::asis_output("\U207A")` T cell populations
```{r annot CD8 T}
NCLS_TIL_FRC_cd8$cell_type <- NULL
NCLS_TIL_FRC_cd8$cell_type <--1

NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 0)] <- paste0("Stem-like/Naive CD8", "\u207A ", "T cells")
NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 1)] <-paste0("Stem-like/Naive CD8", "\u207A ", "T cells")
NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 2)] <- paste0("Effector-memory CD8", "\u207A ", "T cells")
NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 3)] <- paste0("Effector-memory CD8", "\u207A ", "T cells")
NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 4)] <-  paste0("Exhausted CD8", "\u207A ", "T cells")
NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 5)] <- paste0("Effector CD8", "\u207A ", "T cells")
NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 6)] <-paste0("Exhausted CD8", "\u207A ", "T cells")
NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 7)] <- paste0("Exhausted CD8", "\u207A ", "T cells")
NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 8)] <- paste0("Cycling CD8", "\u207A ", "T cells")
NCLS_TIL_FRC_cd8$cell_type[which(NCLS_TIL_FRC_cd8$RNA_snn_res.0.4 == 9)] <- paste0("Effector CD8", "\u207A ", "T cells")
```


### CD8`r knitr::asis_output("\U207A")` T cells in NSCLC (Figure 3D)
```{r umap CD8 T}
#Extend palet for CD8 T cell subsets
palet <- c("#1B9E77", "#54B0E4","#E3BE00", "#E41A1C","#F8766D","#00C08B","#7CAE00","#00B4F0","#F564E3")
names(palet) <- c("TRC","PRC","AdvFB" ,"SMC/PC",paste0("Stem-like/Naive CD8", "\u207A ", "T cells"),paste0("Exhausted CD8", "\u207A ", "T cells"), paste0("Effector-memory CD8", "\u207A ", "T cells"),paste0("Effector CD8", "\u207A ", "T cells"),paste0("Cycling CD8", "\u207A ", "T cells"))

palet <- palet[names(palet) %in% unique(NCLS_TIL_FRC_cd8$cell_type)]

DimPlot(NCLS_TIL_FRC_cd8, reduction = "umap", group.by = "cell_type", cols = palet)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle(paste0("CD8", "\U207A ", "T cells (NSCLC)"))

```

### Save CD8`r knitr::asis_output("\U207A")` T cell data
```{r save CD8 T}
saveRDS(NCLS_TIL_FRC_cd8,paste0(basedir,"/data/Human/NSCLC_TILs_CD8_pop.rds"))
```

### Dotplot CD8`r knitr::asis_output("\U207A")` T cells (Supplementary Figure 3F)
```{r dotplot CD8 T}
genes <-c("CCR7","SELL","TCF7","NKG7","GZMA","GZMB","GZMM","GZMK","CX3CR1","GNLY","FCGR3A",  "KIR3DL2","KLRF1","TIGIT","CTLA4","LAG3","ZEB2","TOP2A","MKI67","PCLAF")

data_conv <-Remove_ensebl_id(NCLS_TIL_FRC_cd8) 
data_conv$cell_type <- factor(data_conv$cell_type, levels = rev(c(paste0("Cycling CD8", "\u207A ", "T cells"),
                                                      paste0("Effector CD8", "\u207A ", "T cells"),
                                                      paste0("Exhausted CD8", "\u207A ", "T cells"),
                                                      paste0("Effector-memory CD8", "\u207A ", "T cells"),
                                                      paste0("Stem-like/Naive CD8", "\u207A ", "T cells"))))
  
gg <-dittoDotPlot(data_conv, vars = genes, group.by = "cell_type", size = 4)
gg + coord_fixed(ratio=0.8) + ylab("")
```

### Session info
```{r session info}
sessionInfo()
date()
```
