---
title: "Ccl19-expressing FRCs control anti-tumor T cell responses"
author: "Chrysa Papadopoulou"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: yes
    collapsed: no
    smooth_scroll: yes
mainfont: {Arial Unicode MS}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("ragg_png", "cairo_pdf"))
knitr::asis_output("\U207A","\U2212")
knitr::spin_child('extra_functions.R')
set.seed(1000)
```

### Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(here)
  library(purrr)
  library(dplyr)
  library(stringr)
  library(patchwork)
  library(Seurat)
  library(Matrix)
  library(gridExtra)
  library(gsubfn)
  library(ggsci)
  library(biomaRt)
  library(tidyverse)
  library(msigdbr)
  library(stats)
  library(clusterProfiler)
  library(dict)
  library(openxlsx)
  library(DOSE)
  library(enrichplot)
  library(dittoSeq)
})
```

### Set directory
```{r set directory}
basedir <- here()
```
 

### Read GP33/34`r knitr::asis_output("\U207A")` CD8`r knitr::asis_output("\U207A")` T cells sorted from LLC-gp33 bearing DTR`r knitr::asis_output("\U207A")` and DTR$^`r knitr::asis_output("\U2212")`$ lungs on day 23, after mCOV-Flt3l-gp33 immunization 
```{r read CD8 T cells}
Tetra_CD8 <- readRDS(paste0(basedir,"/data/Mouse/Tetra_CD8_EXH.rds"))
```

### Set color palette
```{r color vec}
cols <- c("#F8766D","#B79F00","#00C08B","#F564E3","#00B4F0")
names(cols) <-c("Cycling T cells","Eff. Mem. T cells","Effector T cells",paste0("Ccr7", "\U207A ","T cells"),"Exhausted T cells")
```

### CD8`r knitr::asis_output("\U207A")`GP33/34`r knitr::asis_output("\U207A")` T cells (mCOV-FIt31-gp33) (Figure 7F){.tabset}
#### Umap
```{r umap Tetra CD8}
DimPlot(Tetra_CD8, reduction = "umap", group.by = "annot", cols = cols)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle(paste0("CD8", "\U207A ", "GP33/34" , "\U207A ", "T cells (mCOV-FIt31-g33)"))
```

#### DTR`r knitr::asis_output("<sup>\U2212")`
```{r DTR neg umap}
DTRneg <-subset(Tetra_CD8, depleted==paste0("DTR", expression("\u2212")))

DimPlot(DTRneg, reduction = "umap", group.by = "depleted",split.by = "depleted")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

#### DTR$^`r knitr::asis_output("\U207A")`$
```{r DTR pos umap}
DTRpos <-subset(Tetra_CD8, depleted==paste0("DTR", expression("\u207A")))

DimPlot(DTRpos, reduction = "umap", group.by = "depleted",split.by = "depleted")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### Dotplot CD8`r knitr::asis_output("\U207A")` T cell subsets (Supplementary Figure 11E)
```{r dotplot CD8 T cells}
data_conv <-Tetra_CD8
data_conv <-Remove_ensebl_id(data_conv)

Idents(data_conv) <- data_conv$annot
levels(data_conv)<-levels(data_conv)[order(match(levels(data_conv),c(paste0("Ccr7", "\U207A ","T cells"),"Exhausted T cells","Effector T cells","Eff. Mem. T cells","Cycling T cells")))]
data_conv$cell_type <- factor(as.character(data_conv@active.ident), levels = rev(c(paste0("Ccr7", "\U207A ","T cells"),"Exhausted T cells","Effector T cells","Eff. Mem. T cells","Cycling T cells")))

gene_list <-c("Pclaf","Top2a","Mki67","Mcm7","Klrg1","Gzma","Lef1","Klf2","Ifng","Ccl4","Xcl1","Il2","Gzmb","Havcr2","Ctla4","Prf1",
              "Lag3","Pdcd1","Tigit", "Tox","Ccr7","Tcf7","Tnfsf4","Cd83")

dittoDotPlot(data_conv, vars = gene_list, group.by = "cell_type", size = 5,legend.size.title = "Expression (%)",scale = TRUE,summary.fxn.color = mean) + ylab( " ")
```

### Pathway analysis 
#### Enriched in DTR$^`r knitr::asis_output("\U2212")`$ exhausted T cells (Figure 7G)
```{r Tetra Exhausted T}
# Step 1 : Set output directory
subDir <- "GSEA_EXH_T/"
saving_path <- paste0(basedir,"/output/")
final_dir <- file.path(saving_path, subDir)
dir.create(final_dir, showWarnings = FALSE,recursive = TRUE)
map_df <- ExtractMouseGeneSets(final_dir)

# Step 2: Customize parameters
httr::set_config(httr::config(ssl_verifypeer = FALSE))
organism <- "org.Mm.eg.db"

disease_phase <- "Depl_vs_NDepl"
datatype <- "SYMBOL"

Exh_T <-subset(Tetra_CD8, annot %in% "Exhausted T cells")

Idents(Exh_T) <- Exh_T$depleted
  
DEmarkers <-FindAllMarkers(Exh_T, only.pos=T, logfc.threshold = 0.1,
                           min.pct = 0.1)

Vec <-unique(Exh_T$depleted)
EnrichParameters_EXH_T <-customize_parameters(Vec,DEmarkers,organism,datatype,disease_phase,saving_path) 

#Step 3: Plot gene-concept network
pathways <-c("lymphocyte mediated immunity","leukocyte cell-cell adhesion","leukocyte mediated cytotoxicity")

cnetplot(EnrichParameters_EXH_T$enrichcl_list[[2]], node_label="gene", layout = "kk", showCategory = pathways,
        max.overlaps=Inf,color.params = list(gene ="black",
                                             category = "red",
                                             edge = TRUE),cex.params = list(category_label = 0.0000001,
                                            label_gene = 0.000001, gene_label= 0.6)) + 
                                    theme(legend.text = element_text(size = 8))

```

#### Enriched in DTR$^`r knitr::asis_output("\U2212")`$ cycling T cells (Figure 7H)
```{r Tetra cycling T}
# Step 1 : Set output directory
subDir <- "GSEA_CYCL_T/"
saving_path <- paste0(basedir,"/output/")
final_dir <- file.path(saving_path, subDir)
dir.create(final_dir, showWarnings = FALSE,recursive = TRUE)
map_df <- ExtractMouseGeneSets(final_dir)

# Step 2: Customize parameters
httr::set_config(httr::config(ssl_verifypeer = FALSE))
organism <- "org.Mm.eg.db"

disease_phase <- "Depl_vs_NDepl"
datatype <- "SYMBOL"

Cycling_T <-subset(Tetra_CD8, annot %in% "Cycling T cells")

Idents(Cycling_T) <- Cycling_T$depleted
  
DEmarkers <-FindAllMarkers(Cycling_T, only.pos=T, logfc.threshold = 0.1,
                           min.pct = 0.1)

Vec <-unique(Cycling_T$depleted)
EnrichParameters_CYCL_T <-customize_parameters(Vec,DEmarkers,organism,datatype,disease_phase,saving_path) 

#Step 3: Plot gene-concept network
pathways <-c("mitotic cell cycle phase transition","regulation of apoptotic signaling pathway","regulation of leukocyte differentiation")

cnetplot(EnrichParameters_CYCL_T$enrichcl_list[[2]], node_label="gene", layout = "kk", showCategory = pathways,
        max.overlaps=Inf,color.params = list(gene ="black",
                                             category = "red",
                                             edge = TRUE),cex.params = list(category_label = 0.0000001,
                                            label_gene = 0.000001, gene_label= 0.6)) + 
                                    theme(legend.text = element_text(size = 8))

```

#### Enriched in DTR$^`r knitr::asis_output("\U2212")`$ Ccr7`r knitr::asis_output("\U207A")` T cells (Supplementary Figure 11F)
```{r Ccr7 T}
# Step 1 : Set output directory
subDir <- "GSEA_CCR7_T/"
saving_path <- paste0(basedir,"/output/")
final_dir <- file.path(saving_path, subDir)
dir.create(final_dir, showWarnings = FALSE,recursive = TRUE)
map_df <- ExtractMouseGeneSets(final_dir)

# Step 2: Customize parameters
httr::set_config(httr::config(ssl_verifypeer = FALSE))
organism <- "org.Mm.eg.db"

disease_phase <- "Depl_vs_NDepl"
datatype <- "SYMBOL"

Ccr7_T <-subset(Tetra_CD8, annot %in% paste0("Ccr7", "\U207A ","T cells"))

Idents(Ccr7_T) <- Ccr7_T$depleted
  
DEmarkers <-FindAllMarkers(Ccr7_T, only.pos=T, logfc.threshold = 0.1,
                           min.pct = 0.1)

Vec <-unique(Ccr7_T$depleted)
EnrichParameters_CCR7_T <-customize_parameters(Vec,DEmarkers,organism,datatype,disease_phase,saving_path) 
```
#### Enriched in Ccl19-EYFP (DTR$^`r knitr::asis_output("\U2212")`$)
```{r barplot}
ccr7_data <-EnrichParameters_CCR7_T$enrichcl_list[[2]]@result
ccr7_data$Color <-rep("#F564E3",nrow(ccr7_data))
ccr7_data$CellType <-rep(paste0("Ccr7", "\U207A ","T cells"),nrow(ccr7_data))

exh_data <-EnrichParameters_EXH_T$enrichcl_list[[2]]@result
exh_data$Color <-rep("#00B4F0",nrow(exh_data))
exh_data$CellType <-rep("Exhausted T cells",nrow(exh_data))

cycl_data <-EnrichParameters_CYCL_T$enrichcl_list[[2]]@result
cycl_data$Color <-rep("#F8766D",nrow(cycl_data))
cycl_data$CellType <-rep("Cycling T cells",nrow(cycl_data))

df_list <-list(ccr7_data,exh_data,cycl_data)
new_df<-do.call("rbind", df_list)

# Filter on the most significant pathways (keep rows where p.adjust<= 0.05)
test <-new_df[new_df$p.adjust <= 0.05,]

pathways <- c("granzyme-mediated programmed cell death signaling pathway","T cell differentiation",
              "cellular response to type II interferon","lymphocyte migration",
              "chromatin remodeling" ,"alpha-beta T cell activation")

BarPlot(test,pathways,cols)

```

### Session info
```{r session info}
sessionInfo()
date()
```
