---
title: "Total CD8 T cells in mCOV-FIt31-gp33"
author: "Chrysa Papadopoulou"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: yes
    collapsed: no
    smooth_scroll: yes
mainfont: {Arial Unicode MS}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("ragg_png", "cairo_pdf"))
knitr::asis_output("\U207A","\U2212")
knitr::spin_child('extra_functions.R')
set.seed(1000)
```

### Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(here)
  library(purrr)
  library(dplyr)
  library(stringr)
  library(patchwork)
  library(Seurat)
  library(Matrix)
  library(gridExtra)
  library(gsubfn)
  library(ggsci)
  library(biomaRt)
  library(tidyverse)
  library(msigdbr)
  library(stats)
  library(clusterProfiler)
  library(dict)
  library(openxlsx)
  library(DOSE)
  library(enrichplot)
  library(dittoSeq)
  library(CellChat)
})
```

### Set directory
```{r set directory}
basedir <- here()
```
 
## **Total CD8 T cells in DTR**

### Read CD8`r knitr::asis_output("\U207A")` T cells in mCOV-FIt31-gp33
```{r read CD8 T cells}
CD8_T <- readRDS(paste0(basedir,"/data/Mouse/CD3_CD8_annot_final.rds"))
```

### Set color palette
```{r color vec}
cluster_palette <- Polychrome::palette36.colors()
names(cluster_palette) <-unique(CD8_T$annot) 
```

### CD8`r knitr::asis_output("\U207A")` T cells mCOV-FIt31-gp33 {.tabset}
#### Supplementary Figure 6G
```{r umap CD8 T cells}
DimPlot(CD8_T, reduction = "umap", group.by = "annot")+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2") + ggtitle(paste0("CD8", "\U207A ", "T cells (mCOV-FIt31-g33)"))
```

#### Dotplot CD8`r knitr::asis_output("\U207A")` T cell subsets (Supplementary Figure 6E)
```{r dotplot CD8 T cells}
data_conv <-CD8_T
data_conv <-Remove_ensebl_id(data_conv)

gene_list <-c("Pclaf","Top2a","Mki67","Klrg1","Gzma","Cxcr3","Gzmk","Ly6a","Ifng","Ccl4","Xcl1","Gzmb","Havcr2","Ctla4","Prf1","Lag3","Pdcd1","Tigit", "Tox","Ccl5","Bcl2","Il7r","Ccr7","Tcf7","Sell")

dittoDotPlot(data_conv, vars = gene_list, group.by = "annot", size = 5,legend.size.title = "Expression (%)",scale = TRUE) + theme(text = element_text(size = 10)) +ylab( " ") 
```

#### Pathway analysis (Supplementary Figure 6I)
```{r Ccl19-EYFP CD8}
# Step 1 : Set output directory
subDir <- "GSEA_CD8_T/"
saving_path <- paste0(basedir,"/output/")
final_dir <- file.path(saving_path, subDir)
dir.create(final_dir, showWarnings = FALSE,recursive = TRUE)
map_df <- ExtractMouseGeneSets(final_dir)

# Step 2: Customize parameters
httr::set_config(httr::config(ssl_verifypeer = FALSE))
organism <- "org.Mm.eg.db"

disease_phase <- "Depl_vs_NDepl"
datatype <- "SYMBOL"


Idents(CD8_T) <- CD8_T$depleted
  
DEmarkers <-FindAllMarkers(CD8_T, only.pos=T, logfc.threshold = 0.25,
                           min.pct = 0.25)

Vec <-unique(CD8_T$depleted)
EnrichParameters_TLS <-customize_parameters(Vec,DEmarkers,organism,datatype,disease_phase,saving_path) 

# Step 3: Enrichment Analysis
for (i in seq(1,length(EnrichParameters_TLS$enrichcl_list))){
  terms<- EnrichParameters_TLS$enrichcl_list[[i]]
  # Filter on the most significant pathways (keep rows where p.adjust<= 0.05)
  terms<- terms@result[terms@result$p.adjust <= 0.05,]
  population <- Vec[i]
  population<- gsub("/", "_", population)
  write.xlsx(terms, paste0(final_dir,"/","GO_Pathways_",population,".xlsx"),row.names = TRUE)
}

#Step 4: Plot enriched pathways
pathways <-c("cytokine-mediated signaling pathway", "canonical Wnt signaling pathway","V(D)J recombination",
             "alpha-beta T cell differentiation", "interleukin-4 production", "alpha-beta T cell activation")

CD8_terms <- EnrichParameters_TLS$enrichcl_list[[2]]@result

selec_pathways <- CD8_terms[CD8_terms$Description %in% pathways,]

selec_pathways$Description <- factor(selec_pathways$Description, levels = rev(pathways))
selec_pathways <- selec_pathways[order(selec_pathways$Description), ]

ggplot(data=selec_pathways, aes(x=Description, y=qscore, fill = analysis)) + xlab(NULL) +
    geom_bar(stat="identity",position="dodge",colour = "black",show.legend = FALSE, width= 0.8, size = 1 ) + coord_flip() +
    scale_y_continuous(expand = expansion(c(0,0)), limits = c(0.0, 3),breaks = c(0,1,2,3)) +
    scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 80)) +
    theme(aspect.ratio = 1.5, legend.justification = "top",
           plot.title = element_text(hjust = 0.5,size = 12,face="bold"),axis.line = element_line(colour = "black"),
           panel.grid.major = element_blank(),
           panel.grid.minor = element_blank(),
           axis.text.x = element_text(angle = 0, vjust = 0.5,colour = "black", size = 10),
           axis.text.y = element_text(angle = 0, vjust = 0.8,colour = "black", size = 10),
           axis.title.y = element_text(size = rel(2), angle = 45),
           axis.title.x = element_text(size = rel(1.5), angle = 0),
           axis.text = element_text(size = 8),
           panel.background = element_blank(), legend.position = "none") +
    scale_fill_manual(values = "dark gray") + ggtitle(paste0("Enriched in Ccl19-EYFP (DTR","\U2212)"))
```

## **Interactome analysis with Cellchat (Suoqin Jin et al., 2021) between TLS TRC, Sulf1`r knitr::asis_output("\U207A")` TRC and Tetramer`r knitr::asis_output("\U207A")` CD8`r knitr::asis_output("\U207A")` T cells**

### Read GP33/34`r knitr::asis_output("\U207A")` CD8`r knitr::asis_output("\U207A")` T cells sorted from LLC-gp33 bearing DTR`r knitr::asis_output("\U207A")` and DTR$^`r knitr::asis_output("\U2212")`$ lungs on day 23, after mCOV-Flt3l-gp33 immunization 
```{r read Tetra CD8 T cells}
Tetra_CD8 <- readRDS(paste0(basedir,"/data/Mouse/Tetra_CD8_EXH.rds"))
```

### Read CCL19-EYFP`r knitr::asis_output("\U207A")` mCOV-FIt31-g33 cell data
```{r read CCL19 FRC mCOV-FIt31-g33 data}
CCL19_EYFP_mCOV <- readRDS(paste0(basedir,"/data/Mouse/mCOV.rds"))
```

### Subset mCOV-FIt31-g33 data on Sulf1`r knitr::asis_output("\U207A")` TRC and TLS TRC
```{r subset on TRC}
TRCs <- subset(CCL19_EYFP_mCOV, annot %in% c(paste0("Sulf1", "\u207A ", "TRC"), "TLS TRC"))
```

### Merge data
```{r merge data}
data_merge <- merge(Tetra_CD8, y = c(TRCs),
             add.cell.ids = c("Tetra_CD8","TRCs"),
             project = "merge_TRC_CD8_T")

#Preprocessing
resolution <- c(0.1, 0.25, 0.4, 0.6,0.8, 1.)
data_merge  <- preprocessing(data_merge,resolution)
```

### Convert seurat object to cellchat object

```{r conv seurat to cellchat}
data_merge$cell_type <-data_merge$annot

cellchat <- Cellchat_Analysis(data_merge)
cellchat <-CellChatDownstreamAnalysis(cellchat,"mouse",thresh = 0.05)
```

### Set color palette
```{r color palet}
palet <-c("#16FF32", "#3283FE", "#FEAF16" , "#B00068" ,"#1CFFCE","#E41A1C","#1B9E77")
names(palet) <-c("Effector T cells",paste0("Ccr7", "\U207A ","T cells"),"Exhausted T cells","Cycling T cells","Eff. Mem. T cells",paste0("Sulf1", "\u207A ", "TRC"), "TLS TRC") 
```

### Interactome analysis (Supplementary Figure 6J)
```{r interactome analysis, message=FALSE}
gg <- netAnalysis_signalingRole_scatter(cellchat,color.use = palet)
gg <- gg + theme(aspect.ratio = 1.3) + ggtitle("mCOV-FIt31-gp33")
gg
```

We can take a look at all significant interactions and involved signaling pathways

### Significant L-R pairs and related signaling genes for all involved signaling pathways
```{r sign LR pairs}
pathways.show.all <- cellchat@netP$pathways
pairLR.CXCL <- extractEnrichedLR(cellchat, signaling = pathways.show.all, geneLR.return = FALSE)
head(pairLR.CXCL,10)
```

### Cell-cell communication mediated by specific ligand-receptor (L-R) pairs {.tabset}
#### Icam1 - (Itgal+/tgb2) (Supplementary Figure 6K)
```{r ICAM1_ITGAL_ITGB2}
palet<-palet[order(match(names(palet),rownames(cellchat@net$count)))]

# Circle plot
netVisual_individual(cellchat, signaling = pathways.show.all, pairLR.use = "ICAM1_ITGAL_ITGB2", color.use = palet,  layout = "circle")
```
#### Vcam1 - (Itga4+Itgb1) (Supplementary Figure 6K) 
```{r ITGA4_ITGB1_VCAM1}
# Circle plot
netVisual_individual(cellchat, signaling = pathways.show.all, pairLR.use = "ITGA4_ITGB1_VCAM1", color.use = palet,  layout = "circle")
```
#### Cxcl16 - Cxcr6 (Supplementary Figure 6L) 
```{r CXCL16_CXCR6}
# Circle plot
netVisual_individual(cellchat, signaling = pathways.show.all, pairLR.use = "CXCL16_CXCR6", color.use = palet,  layout = "circle")
```
#### Tslp - (Il7r+Crlf2) (Supplementary Figure 6L) 
```{r TSLP_IL7R_CRLF2}
# Circle plot
netVisual_individual(cellchat, signaling = pathways.show.all, pairLR.use = "TSLP_IL7R_CRLF2", color.use = palet,  layout = "circle")
```
#### Tgfb1 - (Tgfbr1+Tgfbr2) (Supplementary Figure 6M) 
```{r TGFB1_TGFBR1_TGFBR2}
# Circle plot
netVisual_individual(cellchat, signaling = pathways.show.all, pairLR.use = "TGFB1_TGFBR1_TGFBR2", color.use = palet,  layout = "circle")
```

#### Mif - (Cd74+ Cd44) (Supplementary Figure 6M) 
```{r MIF_CD74_CD44}
# Circle plot
netVisual_individual(cellchat, signaling = pathways.show.all, pairLR.use = "MIF_CD74_CD44", color.use = palet,  layout = "circle")
```

### Session info
```{r session info}
sessionInfo()
date()
```
